---
title: "MODEL_COMPARISON_FIGURE4"
author: "Matthew J Shannon"
date: "03/11/2022"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
library                  (EnhancedVolcano)
library                  (xlsx)
library                  (SCENIC)
library                  (AUCell)
library                  (GENIE3)
library                  (RcisTarget)
gc                       ()
sessionInfo              ()
```

# Load the integrated and model-specific trophoblast Seurat objects

```{r}
# Integrated (all models) trophoblast object
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"

# MPI (in vivo) trophoblast object
load         (file = "~/Desktop/MPI.Rdata")
DefaultAssay (MPI) <- "integrated"

# PD-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/PD_TOrg.Rdata")
DefaultAssay (PD.TOrg) <- "integrated"

# hTSC-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/hTSC_TOrg.Rdata")
DefaultAssay (hTSC.TOrg) <- "integrated"
```

# Figure 4 and Supplemental Figure 4

## Integrated (in vivo + in vitro) CTBs

```{r}
CTBs <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4"))
CTBs
```

```{r, Figure 4A, integrated CTBs}
UMAPPlot (CTBs, label = TRUE, pt.size = 3, cols = c("#E31A1C", "#999999", "#225EA8", "#984EA3")) + NoLegend ()
```

## Subsetting model-specific CTBs

```{r}
# Maternal-Placental Interface (in vivo) cytotrophoblasts (CTBs)
MPI_CTBs <- subset (CTBs, subset = Model == "MPI")
MPI_CTBs

# Patient-derived trophoblast organoid (in vitro) cytotrophoblasts (CTBs)
PD.TOrg_CTBs <- subset (CTBs, subset = Model == "PD.TOrg")
PD.TOrg_CTBs

# hTSC-derived trophoblast organoid (in vitro) cytotrophoblasts (CTBs)
hTSC.TOrg_CTBs <- subset (CTBs, subset = Model == "hTSC.TOrg")
hTSC.TOrg_CTBs
```

## Re-clustered Maternal-Placental Interface (in vivo) CTBs

```{r, Figure 4A, MPI (in vivo) CTB re-clustering}
MPI_CTBs <- FindNeighbors (MPI_CTBs)
MPI_CTBs <- FindClusters  (MPI_CTBs, resolution = 0.250)
MPI_CTBs <- RunUMAP       (MPI_CTBs, reduction  = "pca", dims = 1:30)
MPI_CTBs <- RunTSNE       (MPI_CTBs, reduction  = "pca", dims = 1:30)
MPI_CTBs
```

```{r, Figure 4A, MPI (in vivo) CTB UMAPs}
# Label all trophoblast clusters by cell type
        new.cluster.ids <- c("CTB 1",  # Cytotrophoblast state 1
                             "CTB 2",  # Cytotrophoblast state 2
                             "CTB 4",  # Cytotrophoblast state 4
                             "CTB 3")  # Cytotrophoblast state 3

names (new.cluster.ids) <- levels       (MPI_CTBs)
MPI_CTBs$celltype       <- Idents       (MPI_CTBs)
MPI_CTBs                <- RenameIdents (MPI_CTBs, new.cluster.ids)

          MPI_CTBs[[]]
UMAPPlot (MPI_CTBs, label = TRUE, pt.size = 3, cols = c("#999999", "#E31A1C", "#984EA3", "#225EA8")) + NoLegend ()
```

## Re-clustered Patient-derived trophoblast organoid (in vitro) CTBs

```{r, Figure 4A, PD-TOrg (in vitro) CTB re-clustering}
PD.TOrg_CTBs <- FindNeighbors (PD.TOrg_CTBs)
PD.TOrg_CTBs <- FindClusters  (PD.TOrg_CTBs, resolution = 0.250)
PD.TOrg_CTBs <- RunUMAP       (PD.TOrg_CTBs, reduction  = "pca", dims = 1:30)
PD.TOrg_CTBs <- RunTSNE       (PD.TOrg_CTBs, reduction  = "pca", dims = 1:30)
PD.TOrg_CTBs
```

```{r, Figure 4A, PD-TOrg (in vitro) CTB UMAPs}
# Label all trophoblast clusters by cell type
          new.cluster.ids <- c("CTB 1",  # Cytotrophoblast state 1
                               "CTB 2",  # Cytotrophoblast state 2
                               "CTB 4",  # Cytotrophoblast state 4
                               "CTB 3")  # Cytotrophoblast state 3

names (new.cluster.ids) <- levels       (PD.TOrg_CTBs)
PD.TOrg_CTBs$celltype    <- Idents       (PD.TOrg_CTBs)
PD.TOrg_CTBs             <- RenameIdents (PD.TOrg_CTBs, new.cluster.ids)

          PD.TOrg_CTBs[[]]
UMAPPlot (PD.TOrg_CTBs, label = TRUE, pt.size = 3, cols = c("#999999", "#225EA8", "#984EA3", "#E31A1C")) + NoLegend ()
```

## Re-clustered hTSC-derived trophoblast organoid (in vitro) CTBs

```{r, Figure 4A, hTSC-TOrg (in vitro) CTB re-clustering}
hTSC.TOrg_CTBs <- FindNeighbors (hTSC.TOrg_CTBs)
hTSC.TOrg_CTBs <- FindClusters  (hTSC.TOrg_CTBs, resolution = 0.150)
hTSC.TOrg_CTBs <- RunUMAP       (hTSC.TOrg_CTBs, reduction  = "pca", dims = 1:30)
hTSC.TOrg_CTBs <- RunTSNE       (hTSC.TOrg_CTBs, reduction  = "pca", dims = 1:30)
hTSC.TOrg_CTBs
```

```{r, Figure 4A, hTSC-TOrg (in vitro) CTB UMAPs}
# Label all trophoblast clusters by cell type
        new.cluster.ids <- c("CTB 1",  # Cytotrophoblast state 1
                             "CTB 4",  # Cytotrophoblast state 4
                             "CTB 2")  # Cytotrophoblast state 2

names (new.cluster.ids) <- levels       (hTSC.TOrg_CTBs)
hTSC.TOrg_CTBs$celltype <- Idents       (hTSC.TOrg_CTBs)
hTSC.TOrg_CTBs          <- RenameIdents (hTSC.TOrg_CTBs, new.cluster.ids)

          hTSC.TOrg_CTBs[[]]
UMAPPlot (hTSC.TOrg_CTBs, label = TRUE, pt.size = 3, cols = c("#999999", "#E31A1C", "#984EA3")) + NoLegend ()
```

## SCENIC analyses (run 5 times each)

### Maternal-Placental Interface (in vivo) CTB (repeat 5 times)

```{r}
CTBs    <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4"))
MPI     <- subset (CTBs, subset = Model == "in vivo")
MPI.CTB <- subset (invivo,  downsample = 500)
MPI.CTB

table (Idents (MPI.CTB))
rm    (Trophoblasts)
rm    (MPI)
```

```{r}
DefaultAssay (MPI.CTB) <- 'RNA'

exprMat  <- MPI.CTB@assays$RNA@data
cellInfo <- MPI.CTB@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_MPI_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32)
 
scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r, Edit code for geneFiltering}
geneFiltering <- function (exprMat, scenicOptions,
                           minCountsPerGene = 3*.01*ncol (exprMat),
                           minSamples       = ncol (exprMat)*.01)
{
  # Load options: outFile_genesKept and dbFilePath
  outFile_genesKept <- NULL
  dbFilePath        <- NULL
  if (class (scenicOptions) == "ScenicOptions")
  {
    dbFilePath        <- getDatabases (scenicOptions)[[1]]
    outFile_genesKept <- getIntName   (scenicOptions, "genesKept")
  }else{
    dbFilePath        <- scenicOptions[["dbFilePath"]]
    outFile_genesKept <- scenicOptions[["outFile_genesKept"]]
  }
  if (is.null (dbFilePath)) stop ("dbFilePath")
  
  # Check expression matrix (e.g. not factor)
  if (is.data.frame (exprMat)) 
  {
    supportedClasses <- paste (gsub ("AUCell_buildRankings,", "", methods ("AUCell_buildRankings")), collapse = ", ")
    supportedClasses <- gsub  ("-method", "", supportedClasses)
    
    stop ("'exprMat' should be one of the following classes: ", supportedClasses, 
          "(data.frames are not supported. Please, convert the expression matrix to one of these classes.)")
  }
  if (any (table (rownames (exprMat)) > 1))
    stop ("The rownames (gene id/name) in the expression matrix should be unique.")
  
  # Calculate stats
  nCountsPerGene <- rowSums (exprMat,     na.rm = TRUE)
  nCellsPerGene  <- rowSums (exprMat > 0, na.rm = TRUE)
  
  ## Show info
  message ("Maximum value in the expression matrix: ", max    (exprMat, na.rm = TRUE))
  message ("Ratio of detected vs non-detected: ",      signif (sum (exprMat > 0, na.rm = TRUE) / sum (exprMat == 0, na.rm = TRUE), 2))
  message ("Number of counts (in the dataset units) per gene:")
  print   (summary (nCountsPerGene))
  message ("Number of cells in which each gene is detected:")
  print   (summary (nCellsPerGene))

  ## Filter
  message ("\nNumber of genes left after applying the following filters (sequential):")
  
  # First filter
  # minCountsPerGene <- 3*.01*ncol (exprMat)
  genesLeft_minReads <- names (nCountsPerGene)[which (nCountsPerGene > minCountsPerGene)]
  message ("\t", length (genesLeft_minReads), "\tgenes with counts per gene > ", minCountsPerGene)
  
  # Second filter
  # minSamples       <- ncol (exprMat)*.01
  nCellsPerGene2     <- nCellsPerGene[genesLeft_minReads]
  genesLeft_minCells <- names (nCellsPerGene2)[which(nCellsPerGene2 > minSamples)]
  message ("\t", length (genesLeft_minCells), "\tgenes detected in more than ", minSamples," cells")
  
  # Exclude genes missing from database:
  library (RcisTarget)
  motifRankings                  <- importRankings (dbFilePath) # either one, they should have the same genes
  genesInDatabase                <- colnames       (getRanking (motifRankings))
  genesLeft_minCells_inDatabases <- genesLeft_minCells[which (genesLeft_minCells %in% genesInDatabase)]
  message ("\t", length (genesLeft_minCells_inDatabases), "\tgenes available in RcisTarget database")
  
  genesKept <- genesLeft_minCells_inDatabases
  if (!is.null (outFile_genesKept)){ 
      saveRDS  (genesKept, file=outFile_genesKept)
    if (getSettings (scenicOptions, "verbose")) message ("Gene list saved in ", outFile_genesKept)
  }
  return (genesKept)
}
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- loadInt  (scenicOptions, "aucell_regulonAUC")
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo)%in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans (binaryRegulonActivity[ , cells, drop = FALSE]))
regulonActivity_byCellType_Binarized <- regulonActivity_byCellType_Binarized[ , -c(3,5,6,7,8,9,11)]
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized>minPerc) > 0), ]
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name="Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
regulonActivity_byCellType_Scaled
```

```{r, Table S7A}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_MPI_CTB_regulons_run#.xlsx")
```

### PD-TOrg (in vitro) CTB (repeat 5 times)

```{r}
PD.TOrg.CTB <- subset (CTBs, subset = Model == "PD.TOrg")

table (Idents (PD.TOrg.CTB))
```

```{r}
DefaultAssay (PD.TOrg.CTB) <- 'RNA'

exprMat  <- PD.TOrg.CTB@assays$RNA@data
cellInfo <- PD.TOrg.CTB@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_PD.TOrg_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32) 

scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- loadInt  (scenicOptions, "aucell_regulonAUC")
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo) %in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans(binaryRegulonActivity[ , cells, drop = FALSE]))
regulonActivity_byCellType_Binarized <- regulonActivity_byCellType_Binarized[, -c(3,5,6,7,8,9,11)]
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized > minPerc) > 0), ]
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name="Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
regulonActivity_byCellType_Scaled
```

```{r, Table S7B}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_PD.TOrg_CTB_regulons_run#.xlsx")
```

### hTSC-TOrg (in vitro) CTB (repeat 5 times)

```{r}
hTSC.TOrg.CTB <- subset (CTBs, subset = Model == "hTSC.TOrg")

table (Idents (hTSC.TOrg.CTB))
```

```{r}
DefaultAssay (hTSC.TOrg.CTB) <- 'RNA'

exprMat  <- hTSC.TOrg.CTB@assays$RNA@data
cellInfo <- hTSC.TOrg.CTB@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_hTSC.TOrg_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32) 

scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- loadInt  (scenicOptions, "aucell_regulonAUC")
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo) %in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans (binaryRegulonActivity[ , cells, drop = FALSE]))
regulonActivity_byCellType_Binarized <- regulonActivity_byCellType_Binarized[ , -c(3,5,6,7,8,9,11)]
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized > minPerc) > 0), ]
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name="Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
regulonActivity_byCellType_Scaled
```

```{r, Table S7C}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_hTSC.TOrg_CTB_regulons_run#.xlsx")
```

### Integrated (all models) trophoblasts (repeat 5 times)

```{r}
Troph.sub  <- subset (Trophoblasts, downsample = 500)

table (Idents (Troph.sub))
rm    (Trophoblasts)
```

```{r}
DefaultAssay (Troph.sub) <- 'RNA'

exprMat  <- Troph.sub@assays$RNA@data
cellInfo <- Troph.sub@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_Trophoblast_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32) 

scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo) %in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans (binaryRegulonActivity[ , cells, drop = FALSE]))
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized > minPerc) > 0), ]
```

```{r, Supplemental figure 4, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name = "Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = T, scale = T))
regulonActivity_byCellType_Scaled
```

```{r, Table S7D}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_Integrated.Trophoblasts_regulons_run#.xlsx")
```

## hTSC-derived trophoblast organoid CTBs and TSCs

```{r, Figure 4E hTSC-derived trophoblast organoid (in vitro) CTB and TSC subset}
          CTB_TSC <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4", "TSC"))
hTSC.TOrg_CTB_TSC <- subset (hTSC.TOrg_CTB_TSC, subset = Model == "hTSC.TOrg")
hTSC.TOrg_CTB_TSC
```

```{r, Figure 4E hTSC-derived trophoblast organoid (in vitro) CTB and TSC re-clustering}
#redo find neighbours/clusters
hTSC.TOrg_CTB_TSC <- FindNeighbors (hTSC.TOrg_CTB_TSC)
hTSC.TOrg_CTB_TSC <- FindClusters  (hTSC.TOrg_CTB_TSC, resolution = 0.100)
hTSC.TOrg_CTB_TSC <- RunUMAP       (hTSC.TOrg_CTB_TSC, reduction  = "pca", dims = 1:30)
hTSC.TOrg_CTB_TSC <- RunTSNE       (hTSC.TOrg_CTB_TSC, reduction  = "pca", dims = 1:30)
hTSC.TOrg_CTB_TSC
```

```{r, Figure 4E hTSC-derived trophoblast organoid (in vitro) CTB and TSC UMAP}
# Label all trophoblast clusters by cell type
           new.cluster.ids <- c("TSC",  # hTSC-enriched state
                                "CTBs") # Cytotrophoblasts

names (new.cluster.ids)    <- levels       (hTSC.TOrg_CTB_TSC)
hTSC.TOrg_CTB_TSC$celltype <- Idents       (hTSC.TOrg_CTB_TSC)
hTSC.TOrg_CTB_TSC          <- RenameIdents (hTSC.TOrg_CTB_TSC, new.cluster.ids)

UMAPPlot (hTSC.TOrg_CTB_TSC, label = TRUE, pt.size = 3, cols = c("#AEA200", "darkgrey")) + NoLegend ()
```

## Plotting the SCENIC results

```{r, Figure 4B and 4C regulon transcription factor dot plot}
         Idents (CTBs) <- CTBs$Model
table   (Idents (CTBs))

DotPlot (CTBs,
         features  = c("SRF",
                       "HSF1",
                       "CEBPD",
                       "ESRRA",
                       "CEBPB",
                       "SMARCA4",
                       "ATF1",
                       "TBL1XR1",
                       "FOSL2",
                       "NFIL3",
                       "ETV5",
                       "ATF4",
                       "EZH2",
                       "XBP1",
                       "DDIT3",
                       "EGR1",
                       "E2F4",
                       "FOXP1",
                       "FOXO4",
                       "IRF6",
                       "TFAP2B",
                       "PBX1",
                       "TEAD4"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

```{r, Figure 4C FOXO4 UMAP}
FeaturePlot (Trophoblasts, features = c("FOXO4"), label = FALSE, pt.size = 5, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
```

```{r, Figure 4F driver gene dot plot}
DotPlot (hTSC.TOrg_CTB_TSC,
         features  = c("AC004510.3",
                       "SLC43A2",
                       "COL4A1",
                       "PSG2",
                       "CLIC3",
                       "GCM1",
                       "ZFP42",
                       "VGLL3",
                       "CENPF",
                       "TFPI"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

```{r, Figure 4G regulon transcription factor dot plot analysis}
TB               <- subset (MPI,       idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4"))
PD               <- subset (PD.TOrg,   idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4"))
TS               <- subset (hTSC.TOrg, idents = c("TSC"))
CTB.TSC          <- merge  (x = TB, y = c(PD, TS), add.cell.ids = NULL, merge.data = TRUE)
Idents (CTB.TSC) <- CTB.TSC$Model
```

```{r, Figure 4G regulon transcription factor dot plot}
DotPlot (CTB.TSC,
         features = c("EGR1",
                      "E2F4",
                      "FOXP1",
                      "FOXO4",
                      "IRF6",
                      "TFAP2B",
                      "PBX1",
                      "TEAD4",
                      "SOX15",
                      "SOX13"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

```{r, Figure 4G UMAPs}
FeaturePlot (Trophoblasts, features = c("TEAD4"), label = FALSE, pt.size = 5, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
FeaturePlot (Trophoblasts, features = c("SOX15"), label = FALSE, pt.size = 5, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
FeaturePlot (Trophoblasts, features = c("SOX13"), label = FALSE, pt.size = 5, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
```

# Supplemental Figure 3

## Model-specific CTB Venn-diagram

```{r, Supplemental figure 3A analysis}
DefaultAssay (MPI_CTBs)       <- "RNA"
DefaultAssay (PD.TOrg_CTBs)   <- "RNA"
DefaultAssay (hTSC.TOrg_CTBs) <- "RNA"

            MPI_CTBs.averages <- AverageExpression (MPI_CTBs,       group.by = "nCount_RNA", return.seurat = FALSE)
        PD.TOrg_CTBs.averages <- AverageExpression (PD.TOrg_CTBs,   group.by = "nCount_RNA", return.seurat = FALSE)
      hTSC.TOrg_CTBs.averages <- AverageExpression (hTSC.TOrg_CTBs, group.by = "nCount_RNA", return.seurat = FALSE)

DefaultAssay (MPI_CTBs)       <- "integrated"
DefaultAssay (PD.TOrg_CTBs)   <- "integrated"
DefaultAssay (hTSC.TOrg_CTBs) <- "integrated"

MPI_CTBsgenes       <- rownames (MPI_CTBs.averages$RNA       [rowSums (MPI_CTBs.averages$RNA       []) > 0, ])
PD.TOrg_CTBsgenes   <- rownames (PD.TOrg_CTBs.averages$RNA   [rowSums (PD.TOrg_CTBs.averages$RNA   []) > 0, ])
hTSC.TOrg_CTBsgenes <- rownames (hTSC.TOrg_CTBs.averages$RNA [rowSums (hTSC.TOrg_CTBs.averages$RNA []) > 0, ])

rm (MPI_CTBs.averages)
rm (PD.TOrg_CTBs.averages)
rm (hTSC.TOrg_CTBs.averages)
```

```{r, Supplemental figure 3A}
venn.diagram (x = list (MPI_CTBsgenes,
                        PD.TOrg_CTBsgenes,
                        hTSC.TOrg_CTBsgenes),
              category.names  = c ("MPI", "PD-TOrg", "hTSC-TOrg"),
              filename        = 'Model_CTB_Venn.svg',
              output          = TRUE,
              imagetype       = "png",
              scaled          = FALSE,
              label.col       = "black",
              col             = "black",
              fontfamily      = "sans",
              cat.default.pos = "text",
              margin          = 0.05)
```

```{r, Table S4B}
Venn_CTBs <- get.venn.partitions (x = list (MPI_CTBsgenes,
                                            PD.TOrg_CTBsgenes,
                                            hTSC.TOrg_CTBsgenes),
                                    force.unique  = TRUE,
                                    keep.elements = TRUE,
                                    hierarchical  = FALSE)
Venn_CTBs
Venn_CTBs$..values..

CTB_17235 <- Venn_CTBs$..values..$`1`
CTB_315   <- Venn_CTBs$..values..$`2`
CTB_603   <- Venn_CTBs$..values..$`3`
CTB_330   <- Venn_CTBs$..values..$`4`
CTB_2128  <- Venn_CTBs$..values..$`5`
CTB_681   <- Venn_CTBs$..values..$`6`
CTB_2551  <- Venn_CTBs$..values..$`7`

write.xlsx (CTB_17235, file = "CTB_17235_Venn.xlsx")
write.xlsx (CTB_315,   file = "CTB_315_Venn.xlsx")
write.xlsx (CTB_603,   file = "CTB_603_Venn.xlsx")
write.xlsx (CTB_330,   file = "CTB_330_Venn.xlsx")
write.xlsx (CTB_2128,  file = "CTB_2128_Venn.xlsx")
write.xlsx (CTB_681,   file = "CTB_681_Venn.xlsx")
write.xlsx (CTB_2551,  file = "CTB_2551_Venn.xlsx")
```

## Model-specific CTB Dot plots

```{r, Supplemental figure 3B, Maternal-Placental Interface (in vivo) CTBs}
DotPlot (MPI_CTBs,
         features  = c("MKI67",
                       "EGFR",
                       "ELF5",
                       "TP63",
                       "SPINT1",
                       "ITGA5",
                       "HLA-G",
                       "ERVW-1",
                       "SDC1",
                       "GCM1"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

```{r, Supplemental figure 3B, patient-derived trophoblast organoid (in vitro) CTBs}
DotPlot (PD.TOrg_CTBs,
         features  = c("MKI67",
                       "EGFR",
                       "ELF5",
                       "TP63",
                       "SPINT1",
                       "ITGA5",
                       "HLA-G",
                       "ERVW-1",
                       "SDC1",
                       "GCM1"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

```{r, Supplemental figure 3B, hTSC-derived trophoblast organoid (in vitro) CTBs}
DotPlot (hTSC.TOrg_CTBs,
         features  = c("MKI67",
                       "EGFR",
                       "ELF5",
                       "TP63",
                       "SPINT1",
                       "ITGA5",
                       "HLA-G",
                       "ERVW-1",
                       "SDC1",
                       "GCM1"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

## Re-clustering all Maternal-Placental Interface (in vivo) and PD-TOrg (in vitro) TSC and CTB states for RNA velocity analysis

### MPI (in vivo)

```{r, Supplemental figure 3D Maternal-Placental Interface (in vivo) CTB and TSC subset and re-clustering}
MPI_CTB_TSC <- subset (CTB_TSC, subset = Model == "MPI")
MPI_CTB_TSC

MPI_CTB_TSC <- FindNeighbors (MPI_CTB_TSC)
MPI_CTB_TSC <- FindClusters  (MPI_CTB_TSC, resolution = 0.250)
MPI_CTB_TSC <- RunUMAP       (MPI_CTB_TSC, reduction  = "pca", dims = 1:30)
```

```{r, Supplemental figure 3D Maternal-Placental Interface (in vivo) CTB and TSC cluster labelling}
# Label all trophoblast clusters by cell type
        new.cluster.ids <- c("CTB 1",  # Cytotrophoblast state 1
                             "CTB 2",  # Cytotrophoblast state 2
                             "CTB 4",  # Cytotrophoblast state 4
                             "CTB 3")  # Cytotrophoblast state 3

names (new.cluster.ids) <- levels       (MPI_CTB_TSC)
MPI_CTB_TSC$celltype    <- Idents       (MPI_CTB_TSC)
MPI_CTB_TSC             <- RenameIdents (MPI_CTB_TSC, new.cluster.ids)
```

```{r, Supplemental figure 3D Maternal-Placental Interface (in vivo) CTB and TSC UMAP}
UMAPPlot (MPI_CTB_TSC, label = TRUE, pt.size = 3, cols = c("#E31A1C", "#999999", "#225EA8", "#984EA3")) + NoLegend ()
```

### PD-TOrg (in vitro)

```{r, Supplemental figure 3D patient-derived trophoblast organoid (in vitro) CTB and TSC subset and re-clustering}
PD.TOrg_CTB_TSC <- subset (CTB_TSC, subset = Model == "PD.TOrg")
PD.TOrg_CTB_TSC

PD.TOrg_CTB_TSC <- FindNeighbors (PD.TOrg_CTB_TSC)
PD.TOrg_CTB_TSC <- FindClusters  (PD.TOrg_CTB_TSC, resolution = 0.250)
PD.TOrg_CTB_TSC <- RunUMAP       (PD.TOrg_CTB_TSC, reduction  = "pca", dims = 1:30)
```

```{r, Supplemental figure 3D patient-derived trophoblast organoid (in vitro) CTB and TSC cluster labelling}
# Label all trophoblast clusters by cell type
         new.cluster.ids <- c("CTB 1",  # Cytotrophoblast state 1
                              "CTB 2",  # Cytotrophoblast state 2
                              "CTB 4",  # Cytotrophoblast state 4
                              "CTB 3",  # Cytotrophoblast state 3
                              "TSC" )   # hTSC-enriched state

names (new.cluster.ids)  <- levels       (PD.TOrg_CTB_TSC)
PD.TOrg_CTB_TSC$celltype <- Idents       (PD.TOrg_CTB_TSC)
PD.TOrg_CTB_TSC          <- RenameIdents (PD.TOrg_CTB_TSC, new.cluster.ids)
```

```{r, Supplemental figure 3D patient-derived trophoblast organoid (in vitro) CTB and TSC UMAP}
UMAPPlot (PD.TOrg_CTB_TSC, label = TRUE, pt.size = 3, cols = c("#E31A1C", "#999999", "#225EA8", "#984EA3", "#AEA200")) + NoLegend ()
```
