---
title: "MODEL_COMPARISON_FIGURE6"
author: "Matthew J Shannon"
date: "15/11/2022"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (ggpubr)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
library                  (EnhancedVolcano)
library                  (xlsx)
gc                       ()
sessionInfo              ()
```

# Load the integrated and model-specific trophoblast Seurat objects

```{r}
# Integrated (all models) trophoblast object
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"

# MPI (in vivo) trophoblast object
load         (file = "~/Desktop/MPI.Rdata")
DefaultAssay (MPI) <- "integrated"

# PD-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/PD_TOrg.Rdata")
DefaultAssay (PD.TOrg) <- "integrated"

# hTSC-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/hTSC_TOrg.Rdata")
DefaultAssay (hTSC.TOrg) <- "integrated"
```

# Figure 6

## Villous Differentiation Branch Comparison

```{r, Figure 6A, Integrated (in vivo + in vitro) villous differentiation branch subset}
Villous <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4", "SCTp 1", "SCTp 2"))
```

```{r, Figure 6A, Integrated (in vivo + in vitro) villous differentiation branch UMAP}
UMAPPlot (Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

### Maternal-Placental Interface (in vivo) villous differentiation branch

```{r, Figure 6A, Maternal-Placental Interface (in vivo) villous differentiation branch re-clustering}
MPI_Villous <- subset        (Villous, subset = Model == "MPI")
MPI_Villous <- FindNeighbors (MPI_Villous)
MPI_Villous <- FindClusters  (MPI_Villous, resolution = 0.025)
MPI_Villous <- RunUMAP       (MPI_Villous, reduction  = "pca", dims = 1:30)
MPI_Villous
```

```{r, Figure 6A, Maternal-Placental Interface (in vivo) villous differentiation branch cluster labelling}
        new.cluster.ids <- c("CTB 1",  # Cytotrophoblast   state 1
                             "CTB 2",  # Cytotrophoblast   state 2
                             "CTB 4",  # Cytotrophoblast   state 4
                             "SCTp",   # Syncytiotrophoblast precursor state
                             "CTB 3")  # Cytotrophoblast   state 3

names (new.cluster.ids) <- levels       (MPI_Villous)
MPI_Villous$celltype    <- Idents       (MPI_Villous)
MPI_Villous             <- RenameIdents (MPI_Villous, new.cluster.ids)
```

```{r, Figure 6A, Maternal-Placental Interface (in vivo) villous differentiation branch UMAP}
UMAPPlot (MPI_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

### Patient-derived trophoblast organoid (in vitro) villous differentiation branch

```{r, Figure 6A, Patient-derived trophoblast organoid (in vitro) villous differentiation branch re-clustering}
PD.TOrg_Villous <- subset        (Villous, subset = Model == "PD.TOrg")
PD.TOrg_Villous <- FindNeighbors (PD.TOrg_Villous)
PD.TOrg_Villous <- FindClusters  (PD.TOrg_Villous, resolution = 0.025)
PD.TOrg_Villous <- RunUMAP       (PD.TOrg_Villous, reduction  = "pca", dims = 1:30)
PD.TOrg_Villous
```

```{r, Figure 6A, Patient-derived trophoblast organoid (in vitro) villous differentiation branch cluster labelling}
         new.cluster.ids <- c("CTB 1",  # Cytotrophoblast state 1
                              "CTB 2",  # Cytotrophoblast state 2
                              "CTB 3",  # Cytotrophoblast state 3
                              "SCTp")   # Syncytiotrophoblast precursor state

names (new.cluster.ids)  <- levels       (PD.TOrg_Villous)
PD.TOrg_Villous$celltype <- Idents       (PD.TOrg_Villous)
PD.TOrg_Villous          <- RenameIdents (PD.TOrg_Villous, new.cluster.ids)
```

```{r, Figure 6A, Patient-derived trophoblast organoid (in vitro) villous differentiation branch UMAP}
UMAPPlot (PD.TOrg_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

### hTSC-derived trophoblast organoid (in vitro) villous differentiation branch

```{r, Figure 6A, hTSC-derived trophoblast organoid (in vitro) villous differentiation branch re-clustering}
hTSC.TOrg_Villous <- subset        (Villous, subset = Model == "hTSC.TOrg")
hTSC.TOrg_Villous <- FindNeighbors (hTSC.TOrg_Villous)
hTSC.TOrg_Villous <- FindClusters  (hTSC.TOrg_Villous, resolution = 0.050)
hTSC.TOrg_Villous <- RunUMAP       (hTSC.TOrg_Villous, reduction  = "pca", dims = 1:30)
hTSC.TOrg_Villous
```

```{r, Figure 6A, hTSC-derived trophoblast organoid (in vitro) villous differentiation branch cluster labelling}
           new.cluster.ids <- c("SCTp 1",  # Syncytiotrophoblast precursor state 1
                                "SCTp 3",  # Syncytiotrophoblast precursor state 3
                                "CTB",     # Cytotrophoblast state
                                "SCTp 2")  # Syncytiotrophoblast precursor state 2

names (new.cluster.ids)    <- levels       (hTSC.TOrg_Villous)
hTSC.TOrg_Villous$celltype <- Idents       (hTSC.TOrg_Villous)
hTSC.TOrg_Villous          <- RenameIdents (hTSC.TOrg_Villous, new.cluster.ids)
```

```{r, Figure 6A, hTSC-derived trophoblast organoid (in vitro) villous differentiation branch UMAP}
UMAPPlot (hTSC.TOrg_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

## Villous Differentiation Branch MPI Driver Gene Comparison

```{r, Figure 6B SCTp subset}
# Integrated trophoblast (in vivo + in vitro) SCTp subset
SCTps <- subset (Trophoblasts, idents = c("SCTp 1", "SCTp 2"))
```

```{r, Figure 6B dot plot}
# Comparing Maternal-Placental Interface scVelo-informed gene driver expression across each trophoblast model
Idents  (SCTps) <- SCTps$Model
DotPlot (SCTps,
         features = c("NCAPG",
                      "ADCY7",
                      "RASA1",
                      "SLC13A4",
                      "TSC22D2",
                      "SLC2A3",
                      "F5",
                      "CLIP1",
                      "AMD1",
                      "OLR1",
                      "DHRS9",
                      "BACE2",
                      "CGA",
                      "SH3BP5",
                      "GBP1",
                      "CYP19A1",
                      "SLC22A11",
                      "EFHD1",
                      "NEURL1",
                      "GREM2"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

## Model-specific villous differentiation branch regulon transcription factor dot plots

```{r, Figure 6C, dot plot (top)}
# Comparing Maternal-Placental Interface scVelo-informed gene driver expression across each trophoblast model
DotPlot (SCTps,
         features = c("HES1",
                      "PITX1",
                      "MXD1",
                      "ETV4",
                      "TFAP2A",
                      "TFAP2C",
                      "NFKB2",
                      "ZNF165",
                      "ZNF143",
                      "MAFF",
                      "MAFK",
                      "MAFG",
                      "GMEB1",
                      "CREM",
                      "CEBPA",
                      "NR2F6"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

## State-specific villous differentiation branch regulon transcription factor dot plots

```{r, Figure 6C, dot plot (bottom)}
# Plotting transcription factors identified in Table S7 across each SCTp state
Idents  (SCTps) <- SCTps$Seurat_clusters
DotPlot (SCTps,
         features = c("HES1",
                      "PITX1",
                      "MXD1",
                      "ETV4",
                      "TFAP2A",
                      "TFAP2C",
                      "NFKB2",
                      "ZNF165",
                      "ZNF143",
                      "MAFF",
                      "MAFK",
                      "MAFG",
                      "GMEB1",
                      "CREM",
                      "CEBPA",
                      "NR2F6"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

```{r, Figure 6C, UMAPs}
FeaturePlot (Trophoblasts, features = c("HES1"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
FeaturePlot (Trophoblasts, features = c("PITX1"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
FeaturePlot (Trophoblasts, features = c("MXD1"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
FeaturePlot (Trophoblasts, features = c("MAFF"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
FeaturePlot (Trophoblasts, features = c("MAFK"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
FeaturePlot (Trophoblasts, features = c("MAFG"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
```

## hTSC-derived trophoblast organoid (in vitro) CTBs, SCTp, and TSCs

```{r, Figure 6D Alternate hTSC-derived trophoblast organoid villous differentiation branch subset}
          ALT_Villous <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4", "SCTp 1", "SCTp 2", "TSC"))
hTSC.TOrg_ALT_Villous <- subset (ALT_Villous, subset = Model == "hTSC.TOrg")
hTSC.TOrg_ALT_Villous
```

```{r, Figure 6D, Alternate hTSC-derived trophoblast organoid (in vitro) villous differentiation branch UMAP}
UMAPPlot (hTSC.TOrg_ALT_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

```{r, Figure 6D, Alternate hTSC-derived trophoblast organoid (in vitro) villous differentiation branch re-clustering}
hTSC.TOrg_ALT_Villous <- FindNeighbors (hTSC.TOrg_ALT_Villous)
hTSC.TOrg_ALT_Villous <- FindClusters  (hTSC.TOrg_ALT_Villous, resolution = 0.150)
hTSC.TOrg_ALT_Villous <- RunUMAP       (hTSC.TOrg_ALT_Villous, reduction  = "pca", dims = 1:30)
```

```{r, Figure 6D, Alternate hTSC-derived trophoblast organoid (in vitro) villous differentiation branch cluster labelling}
# Label all trophoblast clusters by cell type
               new.cluster.ids <- c("TSC",   # hTSC-enriched state
                                    "SCTp",  # Syncytiotrophoblast precursor state
                                    "CTB")   # Cytotrophoblast state

names (new.cluster.ids)        <- levels       (hTSC.TOrg_ALT_Villous)
hTSC.TOrg_ALT_Villous$celltype <- Idents       (hTSC.TOrg_ALT_Villous)
hTSC.TOrg_ALT_Villous          <- RenameIdents (hTSC.TOrg_ALT_Villous, new.cluster.ids)
```

```{r, Figure 6D, Alternate hTSC-derived trophoblast organoid (in vitro) villous differentiation branch re-clustered UMAP}
UMAPPlot (hTSC.TOrg_ALT_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

```{r, Figure 6G dot plot}
# Comparing Alternative villous differentiation branch scVelo-informed gene driver expression across each state
DotPlot (hTSC.TOrg_ALT_Villous,
         features = c("C12orf75",
                      "LAMA5",
                      "AVPI1",
                      "HSPG2",
                      "MAFF",
                      "TIMP3",
                      "APOL6",
                      "HK2",
                      "PITX1",
                      "VGLL3",
                      "MXD1",
                      "MAN1A2",
                      "INTS2",
                      "RHOBTB1",
                      "KYNU",
                      "TMSB4X",
                      "TNFRSF1B",
                      "MTMR10",
                      "TMSB10",
                      "ATF3"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

# Supplemental Figure 6

## Model-specific SCTp comparisons

```{r, Supplemental figure 6A barplot}
ggplot             (SCTps@meta.data, aes (x = c(SCTps$Model), fill = SCTps$seurat_clusters)) +
geom_bar           (position     = "stack") +
theme_classic2     () +
theme              (text         = element_text (size  = 35)) +
theme              (axis.text    = element_text (size  = 35)) +
theme              (axis.title.x = element_text (face  = "bold")) +
theme              (axis.title.y = element_text (face  = "bold")) +
theme              (legend.text  = element_text (size  = 15)) +
theme              (axis.text.y  = element_text (angle = 90)) +
theme              (axis.text.y  = element_text (hjust = 0.5)) +
theme              (axis.text.x  = element_text(colour = "black")) +
theme              (axis.text.y  = element_text(colour = "black")) +
xlab               ("Trophoblast Model") +
ylab               ("Cell Proportion")
```

### SCTp Venn-diagram

```{r, Supplemental figure 6B model-specific subsets}
# Maternal-Placental Interface (in vivo) SCTp subset
MPI_SCTp <- subset (SCTps, subset = Model =="MPI")
MPI_SCTp

# Patient-derived trophoblast organoid (in vitro) SCTp subset
PD.TOrg_SCTp <- subset (SCTps, subset = Model =="PD.TOrg")
PD.TOrg_SCTp

# hTSC-derived trophoblast organoid (in vitro) SCTp subset
hTSC.TOrg_SCTp <- subset (SCTps, subset = Model =="hTSC.TOrg")
hTSC.TOrg_SCTp
```

```{r, Supplemental figure 6B analysis}
DefaultAssay (MPI_SCTp)       <- "RNA"
DefaultAssay (PD.TOrg_SCTp)   <- "RNA"
DefaultAssay (hTSC.TOrg_SCTp) <- "RNA"

            MPI_SCTp.averages <- AverageExpression (MPI_SCTp,       group.by = "nCount_RNA", return.seurat = FALSE)
        PD.TOrg_SCTp.averages <- AverageExpression (PD.TOrg_SCTp,   group.by = "nCount_RNA", return.seurat = FALSE)
      hTSC.TOrg_SCTp.averages <- AverageExpression (hTSC.TOrg_SCTp, group.by = "nCount_RNA", return.seurat = FALSE)

DefaultAssay (MPI_SCTp)       <- "integrated"
DefaultAssay (PD.TOrg_SCTp)   <- "integrated"
DefaultAssay (hTSC.TOrg_SCTp) <- "integrated"

MPI_SCTpgenes       <- rownames (MPI_SCTp.averages$RNA       [rowSums (MPI_SCTp.averages$RNA       []) > 0, ])
PD.TOrg_SCTpgenes   <- rownames (PD.TOrg_SCTp.averages$RNA   [rowSums (PD.TOrg_SCTp.averages$RNA   []) > 0, ])
hTSC.TOrg_SCTpgenes <- rownames (hTSC.TOrg_SCTp.averages$RNA [rowSums (hTSC.TOrg_SCTp.averages$RNA []) > 0, ])

rm (MPI_SCTp.averages)
rm (PD.TOrg_SCTp.averages)
rm (hTSC.TOrg_SCTp.averages)
```

```{r, Supplemental figure 6B}
venn.diagram (x = list (MPI_SCTpgenes,
                        PD.TOrg_SCTpgenes,
                        hTSC.TOrg_SCTpgenes),
              category.names  = c("MPI", "PD-TOrg", "hTSC-TOrg"),
              filename        = 'Model_SCTp_Venn.svg',
              output          = TRUE,
              imagetype       = "png",
              scaled          = FALSE,
              label.col       = "black",
              col             = "black",
              fontfamily      = "sans",
              cat.default.pos = "text",
              margin          = 0.05)
```

```{r, Table S4E}
Venn_SCT <- get.venn.partitions (x = list (MPI_SCTpgenes,
                                           PD.TOrg_SCTpgenes,
                                           hTSC.TOrg_SCTpgenes),
                                 force.unique  = TRUE,
                                 keep.elements = TRUE,
                                 hierarchical  = FALSE)
Venn_SCT
Venn_SCT$..values..

SCT_14685 <- Venn_SCT$..values..$`1`
SCT_1160  <- Venn_SCT$..values..$`2`
SCT_2106  <- Venn_SCT$..values..$`3`
SCT_3459  <- Venn_SCT$..values..$`4`
SCT_102   <- Venn_SCT$..values..$`5`
SCT_176   <- Venn_SCT$..values..$`6`
SCT_722   <- Venn_SCT$..values..$`7`

write.xlsx (SCT_14685, file = "SCT_14685_Venn.xlsx")
write.xlsx (SCT_1160,  file = "SCT_1160_Venn.xlsx")
write.xlsx (SCT_2106,  file = "SCT_2106_Venn.xlsx")
write.xlsx (SCT_3459,  file = "SCT_3459_Venn.xlsx")
write.xlsx (SCT_102,   file = "SCT_102_Venn.xlsx")
write.xlsx (SCT_176,   file = "SCT_176_Venn.xlsx")
write.xlsx (SCT_722,   file = "SCT_722_Venn.xlsx")
```

### SCTp gene expression comparison

```{r, Supplemental figure 6C Dot plot}
DotPlot (SCTp,
         features = c("ERVW-1",
                      "ERVV-1",
                      "ERVFRD-1",
                      "SDC1",
                      "CGA",
                      "SLC40A1",
                      "PAGE4",
                      "SLC52A1",
                      "KRT23",
                      "SLC13A4",
                      "PEG10",
                      "ID2",
                      "DUSP9",
                      "PHLDA2",
                      "SLC26A2",
                      "PSG6",
                      "PSG3",
                      "PSG2",
                      "PSG5",
                      "HIST1H1C",
                      "TFPI2",
                      "LEP",
                      "MT2A",
                      "CGB",
                      "KISS1"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15) 
```

```{r, Supplemental figure 6D SCTp state 1 subset}
        SCTp1 <- subset (Trophoblasts, idents = "SCTp 1")
Idents (SCTp1) <- SCTp1$Model
av.exp         <- AverageExpression (SCTp1)$RNA
cor.exp        <- as.data.frame     (cor (av.exp))
cor.exp$x      <- rownames          (cor.exp)
cor.df         <- tidyr::gather     (data = cor.exp, y, correlation, c("MPI", "PD.TOrg", "hTSC.TOrg"))
cor.df

ggplot               (cor.df, aes (x, y, fill = correlation)) +
geom_tile            () +
scale_fill_gradientn (limits = c(0,1), colours = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100))
```

## Model-specific alternative villous differentiation branch analyses

### Maternal-Placental Interface (in vivo) CTBs, SCTp, and TSCs

```{r, Supplemental figure 6E, Alternate Maternal-Placental Interface (in vivo) villous differentiation branch subset}
MPI_ALT_Villous <- subset (ALT_Villous, subset = Model == "MPI")
MPI_ALT_Villous
```

```{r, Supplemental figure 6E, Alternate Maternal-Placental Interface (in vivo) villous differentiation branch re-clustering}
MPI_ALT_Villous <- FindNeighbors (MPI_ALT_Villous)
MPI_ALT_Villous <- FindClusters  (MPI_ALT_Villous, resolution = 0.250)
MPI_ALT_Villous <- RunUMAP       (MPI_ALT_Villous, reduction  = "pca", dims = 1:30)
MPI_ALT_Villous
```

```{r, Supplemental figure 6E, Alternate Maternal-Placental Interface (in vivo) villous differentiation branch cluster labelling}
         new.cluster.ids <- c("CTB 1",  # Cytotrophoblast state 1
                              "CTB 2",  # Cytotrophoblast state 2
                              "CTB 4",  # Cytotrophoblast state 4
                              "CTB 3",  # Cytotrophoblast state 3
                              "SCTp")   # Syncytiotrophoblast precursor state

names (new.cluster.ids)  <- levels       (MPI_ALT_Villous)
MPI_ALT_Villous$celltype <- Idents       (MPI_ALT_Villous)
MPI_ALT_Villous          <- RenameIdents (MPI_ALT_Villous, new.cluster.ids)
```

```{r, Supplemental figure 6E, Alternate Maternal-Placental Interface (in vivo) villous differentiation branch re-clustered UMAP}
UMAPPlot (MPI_ALT_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

### Patient-derived trophoblast organoid (in vitro) CTBs, SCTp, and TSCs

```{r, Supplemental figure 6E, Alternate patient-derived trophoblast organoid (in vitro) villous differentiation branch subset}
PD.TOrg_ALT_Villous <- subset (ALT_Villous, subset = Model == "PD.TOrg")
PD.TOrg_ALT_Villous
```

```{r, Supplemental figure 6E, Alternate patient-derived trophoblast organoid (in vitro) villous differentiation branch re-clustering}
PD.TOrg_ALT_Villous <- FindNeighbors (PD.TOrg_ALT_Villous)
PD.TOrg_ALT_Villous <- FindClusters  (PD.TOrg_ALT_Villous, resolution = 0.100)
PD.TOrg_ALT_Villous <- RunUMAP       (PD.TOrg_ALT_Villous, reduction  = "pca", dims = 1:30)
PD.TOrg_ALT_Villous
```

```{r, Supplemental figure 6E, Alternate patient-derived trophoblast organoid (in vitro) villous differentiation branch cluster labelling}
             new.cluster.ids <- c("CTB 2",  # Cytotrophoblast state 2
                                  "CTB 1",  # Cytotrophoblast state 1
                                  "CTB 4",  # Cytotrophoblast state 4
                                  "SCTp",   # Syncytiotrophoblast precursor state
                                  "TSC")    # hTSC-enriched state

names (new.cluster.ids)      <- levels       (PD.TOrg_ALT_Villous)
PD.TOrg_ALT_Villous$celltype <- Idents       (PD.TOrg_ALT_Villous)
PD.TOrg_ALT_Villous          <- RenameIdents (PD.TOrg_ALT_Villous, new.cluster.ids)
```

```{r, Supplemental figure 6E, Alternate patient-derived trophoblast organoid (in vitro) villous differentiation branch re-clustered UMAP}
UMAPPlot (PD.TOrg_ALT_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

## Chimeric Maternal-Placental Interface (in vivo) + hTSC-derived trophoblast organoid (in vitro) villous differentiation branch analysis

```{r, Figure 6F, Chimeric (MPI trophoblasts in vivo + in vitro hTSC-TOrg trophoblasts) villous differentiation branch subset and re-clustering}
Chimeric_Villous <- subset        (ALT_Villous, subset = Model == c("MPI", "hTSC.TOrg"))
Chimeric_Villous <- FindNeighbors (Chimeric_Villous)
Chimeric_Villous <- FindClusters  (Chimeric_Villous, resolution = 0.450)
Chimeric_Villous <- RunUMAP       (Chimeric_Villous, reduction  = "pca", dims = 1:30)
```

```{r, Figure 6F, Chimeric (MPI trophoblasts in vivo + in vitro hTSC-TOrg trophoblasts) villous differentiation branch cluster labelling}
# Label all trophoblast clusters by cell type
          new.cluster.ids <- c("TSC",     # hTSC-enriched state
                               "CTB 2",   # Cytotrophoblast state 2
                               "CTB 1",   # Cytotrophoblast state 1
                               "CTB 4",   # Cytotrophoblast state 4
                               "SCTp 1",  # Syncytiotrophoblast precursor state 1
                               "CTB 3",   # Cytotrophoblast state 3
                               "SCTp 2")  # Syncytiotrophoblast precursor state 2

names (new.cluster.ids)   <- levels       (Chimeric_Villous)
Chimeric_Villous$celltype <- Idents       (Chimeric_Villous)
Chimeric_Villous          <- RenameIdents (Chimeric_Villous, new.cluster.ids)
```

```{r, Figure 6F, Chimeric (MPI trophoblasts in vivo + in vitro hTSC-TOrg trophoblasts) villous differentiation branch re-clustered UMAP}
UMAPPlot (Chimeric_Villous, label = TRUE, pt.size = 3) + NoLegend ()
```