---
title: "MODEL_COMPARISON_FIGURE5"
author: "Matthew J Shannon"
date: "03/11/2022"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (ggpubr)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
library                  (EnhancedVolcano)
library                  (xlsx)
gc                       ()
sessionInfo              ()
```

# Load the integrated and model-specific trophoblast Seurat objects

```{r}
# Integrated (all models) trophoblast object
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"

# MPI (in vivo) trophoblast object
load         (file = "~/Desktop/MPI.Rdata")
DefaultAssay (MPI) <- "integrated"

# PD-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/PD_TOrg.Rdata")
DefaultAssay (PD.TOrg) <- "integrated"

# hTSC-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/hTSC_TOrg.Rdata")
DefaultAssay (hTSC.TOrg) <- "integrated"
```

# Figure 5

## ITGA2 Comparison

```{r, Figure 5A ITGA2+ UMAPs}
# Maternal-Placental Interface (in vivo) ITGA2+ UMAP
FeaturePlot (MPI, features = c("ITGA2"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()

# Patient-derived trophoblast organoid (in vitro) ITGA2+ UMAP
FeaturePlot (PD.TOrg, features = c("ITGA2"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()

# hTSC-derived trophoblast organoid (in vitro) ITGA2+ UMAP
FeaturePlot (hTSC.TOrg, features = c("ITGA2"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
```

```{r, Figure 5B}
# Maternal-Placental Interface (in vivo) ITGA2+ trophoblast subset
MPI_ITGA2 <- subset (invivo,  subset = ITGA2 > 0.00)
MPI_ITGA2 # note the total number of cells

# Patient-derived trophoblast organoid (in vitro) ITGA2+ trophoblast subset
PD.TOrg_ITGA2 <- subset (exvivo,  subset = ITGA2 > 0.00)
PD.TOrg_ITGA2 # note the total number of cells

# hTSC-derived trophoblast organoid (in vitro) ITGA2+ trophoblast subset
hTSC.TOrg_ITGA2 <- subset (invitro, subset = ITGA2 > 0.00)
hTSC.TOrg_ITGA2 # note the total number of cells
```

```{r, Figure 5C analysis}
DefaultAssay (MPI_ITGA2)       <- "RNA"
DefaultAssay (PD.TOrg_ITGA2)   <- "RNA"
DefaultAssay (hTSC.TOrg_ITGA2) <- "RNA"

            MPI_ITGA2.averages <- AverageExpression (MPI_ITGA2,       group.by = "nCount_RNA", return.seurat = FALSE)
        PD.TOrg_ITGA2.averages <- AverageExpression (PD.TOrg_ITGA2,   group.by = "nCount_RNA", return.seurat = FALSE)
      hTSC.TOrg_ITGA2.averages <- AverageExpression (hTSC.TOrg_ITGA2, group.by = "nCount_RNA", return.seurat = FALSE)

DefaultAssay (MPI_ITGA2)       <- "integrated"
DefaultAssay (PD.TOrg_ITGA2)   <- "integrated"
DefaultAssay (hTSC.TOrg_ITGA2) <- "integrated"

MPI_ITGA2genes       <- rownames (MPI_ITGA2.averages$RNA       [rowSums (MPI_ITGA2.averages$RNA       []) > 0, ])
PD.TOrg_ITGA2genes   <- rownames (PD.TOrg_ITGA2.averages$RNA   [rowSums (PD.TOrg_ITGA2.averages$RNA   []) > 0, ])
hTSC.TOrg_ITGA2genes <- rownames (hTSC.TOrg_ITGA2.averages$RNA [rowSums (hTSC.TOrg_ITGA2.averages$RNA []) > 0, ])

rm (MPI_ITGA2.averages)
rm (PD.TOrg_ITGA2.averages)
rm (hTSC.TOrg_ITGA2.averages)
```

```{r, Figure 5C}
venn.diagram (x = list (MPI_ITGA2genes,
                        PD.TOrg_ITGA2genes,
                        hTSC.TOrg_ITGA2genes),
              category.names  = c("MPI", "PD-TOrg", "hTSC-TOrg"),
              filename        = 'ITGA2_Venn.png',
              output          = TRUE,
              imagetype       = "png",
              scaled          = FALSE,
              label.col       = "black",
              col             = "black",
              fontfamily      = "sans",
              cat.default.pos = "text",
              margin          = 0.05)
```

```{r, Table S4C}
Venn_ITGA2 <- get.venn.partitions (x = list (MPI_ITGA2genes,
                                             PD.TOrg_ITGA2genes,
                                             hTSC.TOrg_ITGA2genes),
                                  force.unique  = TRUE,
                                  keep.elements = TRUE,
                                  hierarchical  = FALSE)
Venn_ITGA2
Venn_ITGA2$..values..

ITGA2_16349 <- Venn_ITGA2$..values..$`1`
ITGA2_2108  <- Venn_ITGA2$..values..$`2`
ITGA2_1104  <- Venn_ITGA2$..values..$`3`
ITGA2_3109  <- Venn_ITGA2$..values..$`4`
ITGA2_105   <- Venn_ITGA2$..values..$`5`
ITGA2_334   <- Venn_ITGA2$..values..$`6`
ITGA2_513   <- Venn_ITGA2$..values..$`7`

write.xlsx (ITGA2_16349, file = "ITGA2_16349_Venn.xlsx")
write.xlsx (ITGA2_2108,  file = "ITGA2_2108_Venn.xlsx")
write.xlsx (ITGA2_1104,  file = "ITGA2_1104_Venn.xlsx")
write.xlsx (ITGA2_3109,  file = "ITGA2_3109_Venn.xlsx")
write.xlsx (ITGA2_105,   file = "ITGA2_105_Venn.xlsx")
write.xlsx (ITGA2_334,   file = "ITGA2_334_Venn.xlsx")
write.xlsx (ITGA2_513,   file = "ITGA2_513_Venn.xlsx")
```

## Extravillous Differentiation Branch Comparison

```{r, Figure 5E, Integrated (in vivo + in vitro) extravillous differentiation branch subset}
Extravillous <- subset (Trophoblasts, idents = c("TSC", "cCTB 1", "cCTB 2", "cCTB 3", "EVT"))
```

```{r, Figure 5E, Integrated (in vivo + in vitro) extravillous differentiation branch UMAP}
UMAPPlot (Extravillous, label = TRUE, pt.size = 3) + NoLegend ()
```

### Maternal-Placental Interface (in vivo) extravillous differentiation branch

```{r, Figure 5E, Maternal-Placental Interface (in vivo) extravillous differentiation branch re-clustering}
MPI_Extravillous <- subset        (Extravillous, subset = Model == "MPI")
MPI_Extravillous <- FindNeighbors (MPI_Extravillous)
MPI_Extravillous <- FindClusters  (MPI_Extravillous, resolution = 0.09)
MPI_Extravillous <- RunUMAP       (MPI_Extravillous, reduction  = "pca", dims = 1:30)
MPI_Extravillous
```

```{r, Figure 5E, Maternal-Placental Interface (in vivo) extravillous differentiation branch cluster labelling}
          new.cluster.ids <- c("EVT",          # Extravillous trophoblast state
                               "cCTB 2",       # Column cytotrophoblast state 2
                               "ITGA2+ cCTB",  # ITGA2+ column cytotrophoblast state 1
                               "cCTB 1")       # Column cytotrophoblast state 1

names (new.cluster.ids)   <- levels       (MPI_Extravillous)
MPI_Extravillous$celltype <- Idents       (MPI_Extravillous)
MPI_Extravillous          <- RenameIdents (MPI_Extravillous, new.cluster.ids)
```

```{r, Figure 5E, Maternal-Placental Interface (in vivo) extravillous differentiation branch UMAP}
UMAPPlot (MPI_Extravillous, label = TRUE, pt.size = 3) + NoLegend ()
```

### Patient-derived trophoblast organoid (in vitro) extravillous differentiation branch

```{r, Figure 5E, Patient-derived trophoblast organoid (in vitro) extravillous differentiation branch}
PD.TOrg_Extravillous <- subset        (Extravillous, subset = Model == "PD.TOrg")
PD.TOrg_Extravillous <- FindNeighbors (PD.TOrg_Extravillous)
PD.TOrg_Extravillous <- FindClusters  (PD.TOrg_Extravillous, resolution = 0.075)
PD.TOrg_Extravillous <- RunUMAP       (PD.TOrg_Extravillous, reduction  = "pca", dims = 1:30)
PD.TOrg_Extravillous
```

```{r, Figure 5E, Patient-derived trophoblast organoid (in vitro) extravillous differentiation branch cluster labelling}
              new.cluster.ids <- c("cCTB 3",  # Column cytotrophoblast state 3
                                   "cCTB 1",  # Column cytotrophoblast state 1
                                   "EVT",     # Extravillous trophoblast state
                                   "TSC")     # hTSC-enriched state

names (new.cluster.ids)       <- levels       (PD.TOrg_Extravillous)
PD.TOrg_Extravillous$celltype <- Idents       (PD.TOrg_Extravillous)
PD.TOrg_Extravillous          <- RenameIdents (PD.TOrg_Extravillous, new.cluster.ids)
```

```{r, Figure 5E, Patient-derived trophoblast organoid (in vitro) extravillous differentiation branch UMAP}
UMAPPlot (PD.TOrg_Extravillous, label = TRUE, pt.size = 3) + NoLegend ()
```

### hTSC-derived trophoblast organoid (in vitro) extravillous differentiation branch

```{r, Figure 5E, hTSC-derived trophoblast organoid (in vitro) extravillous differentiation branch}
hTSC.TOrg_Extravillous <- subset        (Extravillous, subset = Model == "hTSC.TOrg")
hTSC.TOrg_Extravillous <- FindNeighbors (hTSC.TOrg_Extravillous)
hTSC.TOrg_Extravillous <- FindClusters  (hTSC.TOrg_Extravillous, resolution = 0.100)
hTSC.TOrg_Extravillous <- RunUMAP       (hTSC.TOrg_Extravillous, reduction  = "pca", dims = 1:30)
hTSC.TOrg_Extravillous
```

```{r, Figure 5E, hTSC-derived trophoblast organoid (in vitro) extravillous differentiation branch cluster labelling}
                new.cluster.ids <- c("TSC",     # hTSC-enriched state
                                     "EVT",     # Extravillous trophoblast state
                                     "cCTB 3")  # Column cytotrophoblast state 3

names (new.cluster.ids)         <- levels       (hTSC.TOrg_Extravillous)
hTSC.TOrg_Extravillous$celltype <- Idents       (hTSC.TOrg_Extravillous)
hTSC.TOrg_Extravillous          <- RenameIdents (hTSC.TOrg_Extravillous, new.cluster.ids)
```

```{r, Figure 5E, hTSC-derived trophoblast organoid (in vitro) extravillous differentiation branch UMAP}
UMAPPlot (hTSC.TOrg_Extravillous, label = TRUE, pt.size = 3) + NoLegend ()
```

## Extravillous Differentiation Branch MPI Driver Gene Comparison

```{r, Figure 5F}
# Comparing Maternal-Placental Interface scVelo-informed gene driver expression across each trophoblast model
Idents  (Extravillous) <- Extravillous$Model
DotPlot (Extravillous,
         features  = c("DGKD",
                       "ENC1",
                       "UTRN",
                       "ITGB6",
                       "EPHA2",
                       "DIO2",
                       "TCF7L2",
                       "SYT8",
                       "COL4A1",
                       "PTPRF",
                       "HSPG2",
                       "ITGA1",
                       "PRKAG2",
                       "FAT2",
                       "MAPRE2"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

## Monocle2 extravillous differentiation pseudotime ordering

```{r}
Extravillous.monocle <- subset              (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4", "TSC", "cCTB 1", "cCTB 2", "cCTB 3", "EVT"))
Extravillous.monocle <- as.CellDataSet      (Extravillous.monocle, assay = "RNA")
Extravillous.monocle <- estimateSizeFactors (Extravillous.monocle)
Extravillous.monocle <- estimateDispersions (Extravillous.monocle)
Extravillous.monocle <- detectGenes         (Extravillous.monocle, min_expr = 0.1)
                        print               (head (fData (Extravillous.monocle)))
                        rm                  (Extravillous.monocle)
```

```{r}
Extravillous.monocle_expressed_genes <- row.names (subset (fData (Extravillous.monocle), num_cells_expressed >= 10))
```

```{r}
ERVFRD_id <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "ERVFRD-1"))
HLAG_id   <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "HLA-G"   ))
ITGA1_id  <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "ITGA1"   ))
CDX2_id   <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "CDX2"    ))
BCAM_id   <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "BCAM"    ))
SPINT1_id <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "SPINT1"  ))
ELF5_id   <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "ELF5"    ))
SDC1_id   <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "SDC1"    ))
GCM1_id   <- row.names (subset (fData (Extravillous.monocle), gene_short_name == "GCM1"    ))

Extravillous.monocle.2  <- newCellTypeHierarchy ()
Extravillous.monocle.2  <- addCellType   (Extravillous.monocle.2,
                                          "EVT",
                                          classify_func = function (x) { x [HLAG_id,    ]  >= 2.00 &
                                                                         x [ITGA1_id,   ]  >= 1.00 &
                                                                         x [GCM1_id,    ]  >= 2.00 &
                                                                         x [SDC1_id,    ]  >= 1.00     })
Extravillous.monocle.2  <- addCellType   (Extravillous.monocle.2,
                                          "CTB",
                                          classify_func = function (x) { x [CDX2_id,    ]  >= 0.50 &
                                                                         x [SPINT1_id,  ]  >= 2.00 &
                                                                         x [ELF5_id,    ]  >= 0.50 &
                                                                         x [BCAM_id,    ]  >= 2.00 &
                                                                         x [ERVFRD_id,  ]  <  1.00 &
                                                                         x [SDC1_id,    ]  <  1.00 &
                                                                         x [HLAG_id,    ]  <  2.00 &
                                                                         x [ITGA1_id,   ]  <  1.00     })

Extravillous.monocle    <- classifyCells (Extravillous.monocle, Extravillous.monocle.2)
```

```{r}
marker_diff              <- markerDiffTable   (Extravillous.monocle[Extravillous.monocle_expressed_genes, ], Extravillous.monocle.2, cores = 8)
semisup_clustering_genes <- row.names         (marker_diff) [order (marker_diff$qval)] [1:500]
Extravillous.monocle     <- setOrderingFilter (Extravillous.monocle, semisup_clustering_genes)

rm (Extravillous.monocle.2)
rm (marker_diff)
rm (semisup_clustering_genes)
```

```{r}
Extravillous.monocle <- reduceDimension (Extravillous.monocle,
                                         max_components = 2,
                                         method         = 'DDRTree',
                                         norm_method    = 'none',
                                         pseudo_expr    = 0)
Extravillous.monocle <- orderCells      (Extravillous.monocle)
Extravillous.monocle <- orderCells      (Extravillous.monocle, root_state = 3)
```

### Model-specific Monocle2 extravillous differentiation trajecories

```{r, Figure 5G Monocle2 model-specific trajectories}
plot_cell_trajectory (Extravillous.monocle, color_by = "seurat_clusters", cell_size = 3) + facet_grid (~Model)
```

### Model-specific extravillous gene expression across Monocle2-informed pseudotime

```{r, Figure 5G Monocle2 model-specific trajectory gene expression}
Extravillous_expressed_genes <- row.names (subset (fData (Extravillous.monocle), num_cells_expressed >= 10))
Extravillous_filtered        <- Extravillous.monocle[Extravillous_expressed_genes, ]
```

```{r, Figure 5G Model-specific UTRN gene expression across Monocle2 pseudotime}
my_genes            <- row.names (subset (fData (Extravillous_filtered), gene_short_name %in% "UTRN"))
Extravillous_subset <- Extravillous_filtered[my_genes, ]

plot_genes_in_pseudotime (Extravillous_subset, color_by = "seurat_clusters") + facet_grid (~Model)
```

```{r, Figure 5G Model-specific TCF7L2 gene expression across Monocle2 pseudotime}
my_genes            <- row.names (subset (fData (Extravillous_filtered), gene_short_name %in% "TCF7L2"))
Extravillous_subset <- Extravillous_filtered[my_genes, ]

plot_genes_in_pseudotime (Extravillous_subset, color_by = "seurat_clusters") + facet_grid (~Model)
```

```{r, Figure 5G Model-specific SYT8 gene expression across Monocle2 pseudotime}
my_genes            <- row.names (subset (fData (Extravillous_filtered), gene_short_name %in% "SYT8"))
Extravillous_subset <- Extravillous_filtered[my_genes, ]

plot_genes_in_pseudotime (Extravillous_subset, color_by = "seurat_clusters") + facet_grid (~Model)
```

```{r, Figure 5G Model-specific ITGA1 gene expression across Monocle2 pseudotime}
my_genes            <- row.names (subset (fData (Extravillous_filtered), gene_short_name %in% "ITGA1"))
Extravillous_subset <- Extravillous_filtered[my_genes, ]

plot_genes_in_pseudotime (Extravillous_subset, color_by = "seurat_clusters") + facet_grid (~Model)
```

## Extravillous differentiation branch regulon transcription factor dot plots

```{r, Figure 5H}
# Plotting transcription factors identified in Table S7
DotPlot (Extravillous,
         features  = c("SRF",
                       "MXI1",
                       "KLF6",
                       "SOX4",
                       "TCF7L2",
                       "BATF3",
                       "ASCL2",
                       "XRCC4",
                       "SNAI1",
                       "MXD4",
                       "MYCN",
                       "FOS",
                       "THRB",
                       "FOSL2",
                       "ELF4",
                       "GTF3A"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

# Supplemental Figure 5

```{r, Supplemental figure 5A}
# Plotting enriched TSC gene markers identified in Table S3
DotPlot (Trophoblasts,
         features  = c("ITGA2",
                       "CD44",
                       "FTL",
                       "TMSB4X",
                       "IL32",
                       "UPP1",
                       "NPPB",
                       "ISG15",
                       "NEAT1",
                       "MALAT1",
                       "GADD45A",
                       "7SK.2",
                       "S100A16",
                       "TUBB2A",
                       "CLU",
                       "IFI27",
                       "CD9",
                       "CD81",
                       "IFITM3",
                       "EPCAM"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

## Extravillous differentiation branch cCTB and EVT Venn-diagrams

### cCTB comparisons

```{r, Supplemental figure 5B subset}
# Integrated trophoblast (in vivo + in vitro) cCTB subset
cCTBs <- subset (Trophoblasts, idents = c("cCTB 1", "cCTB 2", "cCTB 3"))

# Maternal-Placental Interface (in vivo) cCTB subset
MPI_cCTB <- subset (cCTBs, subset = Model =="MPI")
MPI_cCTB

# Patient-derived trophoblast organoid (in vitro) cCTB subset
PD.TOrg_cCTB <- subset (cCTBs, subset = Model =="PD.TOrg")
PD.TOrg_cCTB

# hTSC-derived trophoblast organoid (in vitro) cCTB subset
hTSC.TOrg_cCTB <- subset (cCTBs, subset = Model =="hTSC.TOrg")
hTSC.TOrg_cCTB
```

```{r, Supplemental figure 5B analysis}
DefaultAssay (MPI_cCTB)       <- "RNA"
DefaultAssay (PD.TOrg_cCTB)   <- "RNA"
DefaultAssay (hTSC.TOrg_cCTB) <- "RNA"

            MPI_cCTB.averages <- AverageExpression (MPI_cCTB,       group.by = "nCount_RNA", return.seurat = FALSE)
        PD.TOrg_cCTB.averages <- AverageExpression (PD.TOrg_cCTB,   group.by = "nCount_RNA", return.seurat = FALSE)
      hTSC.TOrg_cCTB.averages <- AverageExpression (hTSC.TOrg_cCTB, group.by = "nCount_RNA", return.seurat = FALSE)

DefaultAssay (MPI_cCTB)       <- "integrated"
DefaultAssay (PD.TOrg_cCTB)   <- "integrated"
DefaultAssay (hTSC.TOrg_cCTB) <- "integrated"

MPI_cCTBgenes       <- rownames (MPI_cCTB.averages$RNA       [rowSums (MPI_cCTB.averages$RNA       []) > 0, ])
PD.TOrg_cCTBgenes   <- rownames (PD.TOrg_cCTB.averages$RNA   [rowSums (PD.TOrg_cCTB.averages$RNA   []) > 0, ])
hTSC.TOrg_cCTBgenes <- rownames (hTSC.TOrg_cCTB.averages$RNA [rowSums (hTSC.TOrg_cCTB.averages$RNA []) > 0, ])

rm (MPI_cCTB.averages)
rm (PD.TOrg_cCTB.averages)
rm (hTSC.TOrg_cCTB.averages)
```

```{r, Supplemental figure 5B}
venn.diagram (x = list (MPI_cCTBgenes,
                        PD.TOrg_cCTBgenes,
                        invitro_cCTBgenes),
              category.names  = c("MPI", "PD-TOrg", "hTSC-TOrg"),
              filename        = 'Model_cCTB_Venn.svg',
              output          = TRUE,
              imagetype       = "png",
              scaled          = FALSE,
              label.col       = "black",
              col             = "black",
              fontfamily      = "sans",
              cat.default.pos = "text",
              margin          = 0.05)
```

```{r, Table S4D}
Venn_cCTB <- get.venn.partitions (x = list (MPI_cCTBgenes,
                                            PD.TOrg_cCTBgenes,
                                            hTSC.TOrg_cCTBgenes),
                                 force.unique  = TRUE,
                                 keep.elements = TRUE,
                                 hierarchical  = FALSE)
Venn_cCTB
Venn_cCTB$..values..

cCTB_18193 <- Venn_cCTB$..values..$`1`
cCTB_1387  <- Venn_cCTB$..values..$`2`
cCTB_1548  <- Venn_cCTB$..values..$`3`
cCTB_2526  <- Venn_cCTB$..values..$`4`
cCTB_183   <- Venn_cCTB$..values..$`5`
cCTB_264   <- Venn_cCTB$..values..$`6`
cCTB_874   <- Venn_cCTB$..values..$`7`

write.xlsx (cCTB_18193, file = "cCTB_18193_Venn.xlsx")
write.xlsx (cCTB_1387,  file = "cCTB_1387_Venn.xlsx")
write.xlsx (cCTB_1548,  file = "cCTB_1548_Venn.xlsx")
write.xlsx (cCTB_2526,  file = "cCTB_2526_Venn.xlsx")
write.xlsx (cCTB_183,   file = "cCTB_183_Venn.xlsx")
write.xlsx (cCTB_264,   file = "cCTB_264_Venn.xlsx")
write.xlsx (cCTB_874,   file = "cCTB_874_Venn.xlsx")
```

### EVT comparisons

```{r, Supplemental figure 5C subset}
# Integrated trophoblast (in vivo + in vitro) EVT subset
EVTs <- subset (Trophoblasts, idents = c("EVT"))

# Maternal-Placental Interface (in vivo) EVT subset
MPI_EVT <- subset (EVTs, subset = Model == "MPI")
MPI_EVT

# Patient-derived trophoblast organoid (in vitro) EVT subset
PD.TOrg_EVT <- subset (EVTs, subset = Model == "PD.TOrg")
PD.TOrg_EVT

# hTSC-derived trophoblast organoid (in vitro) EVT subset
hTSC.TOrg_EVT <- subset (EVTs, subset = Model == "hTSC.TOrg")
hTSC.TOrg_EVT
```

```{r, Supplemental figure 5C analysis}
DefaultAssay (MPI_EVT)       <- "RNA"
DefaultAssay (PD.TOrg_EVT)   <- "RNA"
DefaultAssay (hTSC.TOrg_EVT) <- "RNA"

            MPI_EVT.averages <- AverageExpression (MPI_EVT,       group.by = "nCount_RNA", return.seurat = FALSE)
        PD.TOrg_EVT.averages <- AverageExpression (PD.TOrg_EVT,   group.by = "nCount_RNA", return.seurat = FALSE)
      hTSC.TOrg_EVT.averages <- AverageExpression (hTSC.TOrg_EVT, group.by = "nCount_RNA", return.seurat = FALSE)

DefaultAssay (MPI_EVT)       <- "integrated"
DefaultAssay (PD.TOrg_EVT)   <- "integrated"
DefaultAssay (hTSC.TOrg_EVT) <- "integrated"

MPI_EVTgenes       <- rownames (MPI_EVT.averages$RNA       [rowSums (MPI_EVT.averages$RNA       []) > 0, ])
PD.TOrg_EVTgenes   <- rownames (PD.TOrg_EVT.averages$RNA   [rowSums (PD.TOrg_EVT.averages$RNA   []) > 0, ])
hTSC.TOrg_EVTgenes <- rownames (hTSC.TOrg_EVT.averages$RNA [rowSums (hTSC.TOrg_EVT.averages$RNA []) > 0, ])

rm  (MPI_EVT.averages)
rm  (PD.TOrg_EVT.averages)
rm (hTSC.TOrg_EVT.averages)
```

```{r, Supplemental figure 5C}
venn.diagram (x = list (MPI_EVTgenes,
                        PD.TOrg_EVTgenes,
                        hTSC.TOrg_EVTgenes),
              category.names  = c("MPI", "PD-TOrg", "hTSC-TOrg"),
              filename        = 'Model_EVT_Venn.svg',
              output          = TRUE,
              imagetype       = "png",
              scaled          = FALSE,
              label.col       = "black",
              col             = "black",
              fontfamily      = "sans",
              cat.default.pos = "text",
              margin          = 0.05)
```

```{r, Table S4E}
Venn_EVT <- get.venn.partitions (x = list (MPI_EVTgenes,
                                           PD.TOrg_EVTgenes,
                                           hTSC.TOrg_EVTgenes),
                                 force.unique  = TRUE,
                                 keep.elements = TRUE,
                                 hierarchical  = FALSE)
Venn_EVT
Venn_EVT$..values..

EVT_15031 <- Venn_EVT$..values..$`1`
EVT_567   <- Venn_EVT$..values..$`2`
EVT_2064  <- Venn_EVT$..values..$`3`
EVT_1657  <- Venn_EVT$..values..$`4`
EVT_286   <- Venn_EVT$..values..$`5`
EVT_270   <- Venn_EVT$..values..$`6`
EVT_1629  <- Venn_EVT$..values..$`7`

write.xlsx (EVT_15031, file = "EVT_15031_Venn.xlsx")
write.xlsx (EVT_567,   file = "EVT_567_Venn.xlsx")
write.xlsx (EVT_2064,  file = "EVT_2064_Venn.xlsx")
write.xlsx (EVT_1657,  file = "EVT_1657_Venn.xlsx")
write.xlsx (EVT_286,   file = "EVT_286_Venn.xlsx")
write.xlsx (EVT_270,   file = "EVT_270_Venn.xlsx")
write.xlsx (EVT_1629,  file = "EVT_1629_Venn.xlsx")
```

## EVT re-clustering

```{r, Supplemental figure 5D EVT re-clustering}
EVTs <- FindNeighbors (EVTs)
EVTs <- FindClusters  (EVTs, resolution = 0.100)
EVTs <- RunUMAP       (EVTs, reduction  = "pca", dims = 1:30)
```

```{r, Supplemental figure 5D EVT cluster labelling}
# Label all trophoblast clusters by cell type
        new.cluster.ids <- c("EVT 1",   # Extravillous trophoblast state 1
                             "EVT 2")   # Extravillous trophoblast state 2

names (new.cluster.ids) <- levels       (EVTs)
EVTs$celltype           <- Idents       (EVTs)
EVTs                    <- RenameIdents (EVTs, new.cluster.ids)
EVTs[[]]
```

```{r, Supplemental figure 5D EVT UMAP}
UMAPPlot (EVTs, label = TRUE, pt.size = 3) + NoLegend ()
```

```{r, Supplemental figure 5D EVT barplot}
ggplot             (EVTs@meta.data, aes (x = c(EVTs$Model), fill = EVTs$seurat_clusters)) + 
geom_bar           (position     = "stack") + 
theme_classic2     () +
theme              (text         = element_text (size  = 35)) + 
theme              (axis.text    = element_text (size  = 35)) + 
theme              (axis.title.x = element_text (face  = "bold")) +
theme              (axis.title.y = element_text (face  = "bold")) + 
theme              (legend.text  = element_text (size  = 15)) +
theme              (axis.text.y  = element_text (angle = 90)) +
theme              (axis.text.y  = element_text (hjust = 0.5)) +
theme              (axis.text.x  = element_text(colour = "black")) +
theme              (axis.text.y  = element_text(colour = "black")) +
xlab               ("Trophoblast Model") +
ylab               ("Cell Proportion") 
NoLegend           ()
```

```{r, Supplemental figure 5E EVT state 1 sbuset and model-specific PCC comparison}
        EVT_state1  <- subset (EVTs, idents = c("EVT 1"))
Idents (EVT_state1) <- EVT_state1$Model
av.exp              <- AverageExpression (EVT_state1)$RNA
cor.exp             <- as.data.frame     (cor (av.exp))
cor.exp$x           <- rownames          (cor.exp)
cor.df              <- tidyr::gather     (data = cor.exp, y, correlation, c("MPI", "PD.TOrg", "hTSC.TOrg"))
cor.df

ggplot               (cor.df, aes (x, y, fill = correlation)) +
geom_tile            () +
scale_fill_gradientn (limits = c(0,1), colours = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100))
```

## Extravillous differentiation branch gene expression across Monocle2-informed pseudotime

```{r, Supplemental figure 5F}
plot_cell_trajectory (Extravillous.monocle, markers = "SPINT1", use_color_gradient = TRUE) + NoLegend () + facet_grid (~Model)
plot_cell_trajectory (Extravillous.monocle, markers = "BCAM",   use_color_gradient = TRUE) + NoLegend () + facet_grid (~Model)
plot_cell_trajectory (Extravillous.monocle, markers = "EPCAM",  use_color_gradient = TRUE) + NoLegend () + facet_grid (~Model)
plot_cell_trajectory (Extravillous.monocle, markers = "ITGA2",  use_color_gradient = TRUE) + NoLegend () + facet_grid (~Model)
plot_cell_trajectory (Extravillous.monocle, markers = "ITGA5",  use_color_gradient = TRUE) + NoLegend () + facet_grid (~Model)
plot_cell_trajectory (Extravillous.monocle, markers = "HLA-G",  use_color_gradient = TRUE) + NoLegend () + facet_grid (~Model)
```
