---
title:  "MODEL-SPECIFIC DATA PREPARATION"
author: "Matthew J Shannon"
date:   "17/10/2022"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
```

# Load the integrated trophoblast Seurat object

```{r}
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"
```

```{r}
UMAPPlot (Trophoblasts, pt.size = 3, label = TRUE, split.by = 'Model') + NoLegend ()
```

# Subset the dataset into model-specific trophoblast groups

```{r}
# Maternal-placental interface (in vivo) trophoblasts
MPI <- subset (Trophoblasts, subset = Model == "MPI")
MPI

# PD-TOrg (in vitro) trophoblasts
PD.TOrg <- subset (Trophoblasts, subset = Model == "PD.TOrg")
PD.TOrg

# hTSC-TOrg (in vitro) trophoblasts
hTSC.TOrg <- subset (Trophoblasts, subset = Model == "hTSC.TOrg")
hTSC.TOrg
```

# Quantify the number of cells within each trophoblast state in each model

```{r, Table S5}
# Table S5B
table (Idents (MPI))

# Table S5C
table (Idents (PD.TOrg))

# Table S5D
table (Idents (hTSC.TOrg))
```

```{r}
rm (Trophoblasts)
```

# Prepare the maternal-placental interface (in vivo) trophoblast object

```{r}
DefaultAssay (MPI) <- "integrated"

#redo find neighbours/clusters
MPI <- FindNeighbors (MPI)
MPI <- FindClusters  (MPI, resolution = 0.375)
# Reduce trophoblast data dimensionality
MPI <- RunUMAP       (MPI, reduction  = "pca", dims = 1:30)
MPI <- RunTSNE       (MPI, reduction  = "pca", dims = 1:30)
MPI
```

```{r}
# Visualize the trophoblast data by TSNE
TSNEPlot    (MPI, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize the trophoblast data by UMAP
UMAPPlot    (MPI, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize expression of trophoblast QC metrics by UMAP
FeaturePlot (MPI, features = c("nFeature_RNA"), min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (MPI, features = c("nCount_RNA"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (MPI, features = c("percent.mt"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
```

```{r}
# Label all clusters
MPI <- RenameIdents (MPI,
                        '0'      = "CTB 1",
                        '1'      = "CTB 4",
                        '2'      = "CTB 2",
                        '3'      = "cCTB 2",
                        '4'      = "EVT",
                        '5'      = "CTB 3",
                        '6'      = "SCTp",
                        '7'      = "cCTB 1")
# Bin states into trophoblast subtypes
MPI <- RenameIdents (MPI,
                        'CTB 1'  = "CTB",
                        'CTB 4'  = "CTB",
                        'CTB 2'  = "CTB",
                        'cCTB 2' = "cCTB",
                        'EVT'    = "EVT",
                        'CTB 3'  = "CTB",
                        'SCTp'   = "SCTp",
                        'cCTB 1' = "cCTB")

MPI$celltype <- Idents (MPI)
MPI[[]]

UMAPPlot (MPI, label = TRUE) + NoLegend ()
```

# Identify enriched maternal-placental interface (in vivo) trophoblast-state gene markers

```{r}
DefaultAssay (MPI) <- "RNA"

MPI.markers <- FindAllMarkers (object   = MPI,
                               only.pos = TRUE,
                               min.pct  = 0.25, 
                               test.use = "MAST")

DefaultAssay (MPI) <- "integrated"
```

```{r}
MPI.markers %>% group_by (cluster) %>% top_n (20, avg_log2FC)
```

```{r, Table S3B}
write.xlsx (MPI.markers, file = "MPI_genemarkers.xlsx")
```

# Save the maternal-placental interface (in vivo) trophoblast object

```{r}
save (MPI, file = "MPI_Trophoblasts.Rdata")
```

# Prepare the PD-TOrg (in vitro) trophoblast object

```{r}
DefaultAssay (PD.TOrg) <- "integrated"

#redo find neighbours/clusters
PD.TOrg <- FindNeighbors (PD.TOrg)
PD.TOrg <- FindClusters  (PD.TOrg, resolution = 0.15)
# Reduce trophoblast data dimensionality
PD.TOrg <- RunUMAP       (PD.TOrg, reduction  = "pca", dims = 1:30)
PD.TOrg <- RunTSNE       (PD.TOrg, reduction  = "pca", dims = 1:30)
PD.TOrg
```

```{r}
# Visualize the trophoblast data by TSNE
TSNEPlot    (PD.TOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize the trophoblast data by UMAP
UMAPPlot    (PD.TOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize expression of trophoblast QC metrics by UMAP
FeaturePlot (PD.TOrg, features = c("nFeature_RNA"), min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (PD.TOrg, features = c("nCount_RNA"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (PD.TOrg, features = c("percent.mt"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
```

```{r}
# Label all clusters
PD.TOrg <- RenameIdents (PD.TOrg,
                        '0'     = "CTB 1",
                        '1'     = "cCTB",
                        '2'     = "CTB 2",
                        '3'     = "EVT",
                        '4'     = "SCTp",
                        '5'     = "PTB")
# Bin states into trophoblast subtypes
PD.TOrg <- RenameIdents (PD.TOrg,
                        'CTB 1' = "CTB",
                        'cCTB'  = "cCTB",
                        'CTB 2' = "CTB",
                        'EVT'   = "EVT",
                        'SCTp'  = "SCTp",
                        'PTB'   = "TSC")

PD.TOrg$celltype <- Idents (PD.TOrg)
PD.TOrg[[]]

UMAPPlot (PD.TOrg, label = TRUE) + NoLegend ()
```

# Identify enriched PD-TOrg (in vitro) trophoblast-state gene markers

```{r}
DefaultAssay (PD.TOrg) <- "RNA"

PD.TOrg.markers <- FindAllMarkers (object   = PD.TOrg,
                                   only.pos = TRUE,
                                   min.pct  = 0.25, 
                                   test.use = "MAST")

DefaultAssay (PD.TOrg) <- "integrated"
```

```{r}
PD.TOrg.markers %>% group_by (cluster) %>% top_n (20, avg_log2FC)
```

```{r, Table S3C}
write.xlsx (PD.TOrg.markers, file = "PD-TOrg_genemarkers.xlsx")
```

# Save the PD-TOrg (in vitro) trophoblast object

```{r}
save (PD.TOrg, file = "PD_TOrg.Rdata")
```

# Prepare the hTSC-TOrg (in vitro) trophoblast object

```{r}
DefaultAssay (hTSC.TOrg) <- "integrated"

#redo find neighbours/clusters
hTSC.TOrg  <- FindNeighbors (hTSC.TOrg)
hTSC.TOrg  <- FindClusters  (hTSC.TOrg, resolution = 0.25)
# Reduce trophoblast data dimensionality
hTSC.TOrg  <- RunUMAP       (hTSC.TOrg, reduction  = "pca", dims = 1:30)
hTSC.TOrg  <- RunTSNE       (hTSC.TOrg, reduction  = "pca", dims = 1:30)
hTSC.TOrg
```

```{r}
# Visualize the trophoblast data by TSNE
TSNEPlot    (hTSC.TOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize the trophoblast data by UMAP
UMAPPlot    (hTSC.TOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize expression of trophoblast QC metrics by UMAP
FeaturePlot (hTSC.TOrg, features = c("nFeature_RNA"), min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (hTSC.TOrg, features = c("nCount_RNA"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (hTSC.TOrg, features = c("percent.mt"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
```

```{r}
# Label all clusters
hTSC.TOrg <- RenameIdents (hTSC.TOrg,
                         '0'      = "TSC 1",
                         '1'      = "cCTB",
                         '2'      = "SCTp 2",
                         '3'      = "CTB",
                         '4'      = "SCTp 1",
                         '5'      = "TSC 2",
                         '6'      = "EVT" )
# Bin states into trophoblast subtypes
hTSC.TOrg <- RenameIdents (hTSC.TOrg,
                         'TSC 1'  = "TSC",
                         'cCTB'   = "cCTB",
                         'SCTp 2' = "SCTp",
                         'CTB'    = "CTB",
                         'SCTp 1' = "SCTp",
                         'TSC 2'  = "TSC",
                         'EVT'    = "EVT" )

hTSC.TOrg$celltype <- Idents (hTSC.TOrg)
hTSC.TOrg[[]]

UMAPPlot (hTSC.TOrg, label = TRUE) + NoLegend ()
```

# Identify enriched hTSC-TOrg (in vitro) trophoblast-state gene markers

```{r}
DefaultAssay (hTSC.TOrg) <- "RNA"

hTSC.TOrg.markers <- FindAllMarkers (object   = hTSC.TOrg,
                                     only.pos = TRUE,
                                     min.pct  = 0.25, 
                                     test.use = "MAST")

DefaultAssay (hTSC.TOrg) <- "integrated"
```

```{r}
hTSC.TOrg.markers %>% group_by (cluster) %>% top_n (20, avg_log2FC)
```

```{r, Table S3D}
write.xlsx (hTSC.TOrg.markers, file = "hTSC-TOrg_genemarkers.xlsx")
```

# Save the hTSC-TOrg (in vitro) trophoblast object

```{r}
save (hTSC.TOrg, file = "hTSC_TOrg.Rdata")
```
