---
title: "MODEL_COMPARISON_FIGURE3"
author: "Matthew J Shannon"
date: "08/11/2022"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
library                  (EnhancedVolcano)
library                  (monocle2)
library                  (monocle3)
library                  (patchwork)
library                  (magrittr)
library                  (ggpubr)
```

# Load the integrated and model-specific trophoblast Seurat objects

```{r}
# Integrated (all models) trophoblast object
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"

# MPI (in vivo) trophoblast object
load         (file = "~/Desktop/MPI.Rdata")
DefaultAssay (MPI) <- "integrated"

# PD-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/PD_TOrg.Rdata")
DefaultAssay (PD.TOrg) <- "integrated"

# hTSC-TOrg (in vitro) trophoblast object
load         (file = "~/Desktop/hTSC_TOrg.Rdata")
DefaultAssay (hTSC.TOrg) <- "integrated"
```

# Supplemental Figure 2

## Model-specific orgin gene signatures

```{r, Supplmental figure 2A}
# Plotting the MPI (in vivo) progenitor gene signature
DefaultAssay (MPI) <- "RNA"
FeaturePlot  (MPI, features = c("TP63"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (MPI, features = c("BCAM"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (MPI, features = c("SPINT1"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (MPI, features = c("TEAD4"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (MPI, features = c("ELF5"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (MPI, features = c("EGFR"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
DefaultAssay (MPI) <- "integrated"

# Plotting the PD-TOrg (in vitro) progenitor gene signature
DefaultAssay (PD.TOrg) <- "RNA"
FeaturePlot  (PD.TOrg, features = c("TP63"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (PD.TOrg, features = c("BCAM"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (PD.TOrg, features = c("SPINT1"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (PD.TOrg, features = c("TEAD4"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (PD.TOrg, features = c("ELF5"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (PD.TOrg, features = c("EGFR"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
DefaultAssay (PD.TOrg) <- "integrated"

# Plotting the hTSC-TOrg (in vitro) progenitor gene signature
DefaultAssay (hTSC.TOrg) <- "RNA"
FeaturePlot  (hTSC.TOrg, features = c("TP63"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (hTSC.TOrg, features = c("BCAM"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (hTSC.TOrg, features = c("SPINT1"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (hTSC.TOrg, features = c("TEAD4"),  label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (hTSC.TOrg, features = c("ELF5"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
FeaturePlot  (hTSC.TOrg, features = c("EGFR"),   label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red"))
DefaultAssay (hTSC.TOrg) <- "integrated"
```

## Model-specific origin selection/subset

### MPI (in vivo) trophoblast origin subset

```{r, Supplemental figure 2B MPI origin subset}
MPI_ORIGIN <- subset (MPI, idents = c("CTB 1", "CTB 3")) # Use the non-binned cell clusters
MPI_ORIGIN
```

```{r, Supplemental figure 2B MPI origin re-clustering}
#redo find neighbours/clusters
MPI_ORIGIN <- FindNeighbors (MPI_ORIGIN)
MPI_ORIGIN <- FindClusters  (MPI_ORIGIN, resolution = 0.250)

# Reduce trophoblast data dimensionality
MPI_ORIGIN <- RunUMAP (MPI_ORIGIN, reduction  = "pca", dims = 1:30)
MPI_ORIGIN <- RunTSNE (MPI_ORIGIN, reduction  = "pca", dims = 1:30)
MPI_ORIGIN
```

```{r, Supplemental figure 2B MPI origin UMAP}
UMAPPlot (MPI_ORIGIN, label = TRUE, pt.size = 3) + NoLegend ()
```

### PD-TOrg (in vitro) trophoblast origin subset

```{r, Supplemental figure 2B PD-TOrg origin subset}
PD.TOrg_ORIGIN <- subset (PD.TOrg, idents = c("CTB 1")) # Use the non-binned cell clusters
PD.TOrg_ORIGIN
```

```{r, Supplemental figure 2B PD-TOrg origin re-clustering}
#redo find neighbours/clusters
PD.TOrg_ORIGIN <- FindNeighbors (PD.TOrg_ORIGIN)
PD.TOrg_ORIGIN <- FindClusters  (PD.TOrg_ORIGIN, resolution = 0.150)

# Reduce trophoblast data dimensionality
PD.TOrg_ORIGIN <- RunUMAP (PD.TOrg_ORIGIN, reduction  = "pca", dims = 1:30)
PD.TOrg_ORIGIN <- RunTSNE (PD.TOrg_ORIGIN, reduction  = "pca", dims = 1:30)
PD.TOrg_ORIGIN
```

```{r, Supplemental figure 2B PD-TOrg origin UMAP}
UMAPPlot (PD.TOrg_ORIGIN, label = TRUE, pt.size = 3) + NoLegend ()
```

### hTSC-TOrg (in vitro) trophoblast origin subset

```{r, Supplemental figure 2B hTSC-TOrg origin subset}
hTSC.TOrg_ORIGIN <- subset (hTSC.TOrg, idents = c("TSC 1")) # Use the non-binned cell clusters
hTSC.TOrg_ORIGIN
```

```{r, Supplemental figure 2B hTSC-TOrg origin re-clustering}
#redo find neighbours/clusters
hTSC.TOrg_ORIGIN <- FindNeighbors (hTSC.TOrg_ORIGIN)
hTSC.TOrg_ORIGIN <- FindClusters  (hTSC.TOrg_ORIGIN, resolution = 0.200)

# Reduce trophoblast data dimensionality
hTSC.TOrg_ORIGIN <- RunUMAP (hTSC.TOrg_ORIGIN, reduction  = "pca", dims = 1:30)
hTSC.TOrg_ORIGIN <- RunTSNE (hTSC.TOrg_ORIGIN, reduction  = "pca", dims = 1:30)
hTSC.TOrg_ORIGIN
```

```{r, Supplemental figure 2B hTSC-TOrg origin UMAP}
UMAPPlot (hTSC.TOrg_ORIGIN, label = TRUE, pt.size = 3) + NoLegend ()
```

# Figure 3

## Monocle2 pseudotime analyses

### MPI (in vivo) trophoblasts

```{r, Figure 3A MPI (in vivo) analysis}
MPI.monocle <- as.CellDataSet      (MPI, assay = "RNA")
MPI.monocle <- estimateSizeFactors (MPI.monocle)
MPI.monocle <- estimateDispersions (MPI.monocle)
MPI.monocle <- detectGenes         (MPI.monocle, min_expr = 0.1)
               print               (head (fData (MPI.monocle)))
```

```{r, Figure 3A MPI (in vivo) analysis}
MPI.monocle_expressed_genes <- row.names (subset (fData (MPI.monocle), num_cells_expressed >= 10))
```

```{r, Figure 3A MPI (in vivo) analysis}
ERVFRD_id <- row.names (subset (fData (MPI.monocle), gene_short_name == "ERVFRD-1"))
HLAG_id   <- row.names (subset (fData (MPI.monocle), gene_short_name == "HLA-G"   ))
CDX2_id   <- row.names (subset (fData (MPI.monocle), gene_short_name == "CDX2"    ))
BCAM_id   <- row.names (subset (fData (MPI.monocle), gene_short_name == "BCAM"    ))
SPINT1_id <- row.names (subset (fData (MPI.monocle), gene_short_name == "SPINT1"  ))
ELF5_id   <- row.names (subset (fData (MPI.monocle), gene_short_name == "ELF5"    ))
SDC1_id   <- row.names (subset (fData (MPI.monocle), gene_short_name == "SDC1"    ))
GCM1_id   <- row.names (subset (fData (MPI.monocle), gene_short_name == "GCM1"    ))

MPI.monocle.2 <- newCellTypeHierarchy ()
MPI.monocle.2 <- addCellType   (MPI.monocle.2,
                                "SCT",
                                classify_func = function (x) { x [ERVFRD_id,  ]  >= 1.00 &
                                                               x [GCM1_id,    ]  >= 2.00 &
                                                               x [SDC1_id,    ]  >= 1.00   })
MPI.monocle.2 <- addCellType   (MPI.monocle.2,
                                "EVT",
                                classify_func = function (x) { x [HLAG_id,    ]  >= 2.00 &
                                                               x [GCM1_id,    ]  >= 2.00 &
                                                               x [SDC1_id,    ]  >= 1.00   })
MPI.monocle.2 <- addCellType   (MPI.monocle.2,
                                "CTB",
                                classify_func = function (x) { x [CDX2_id,    ]  >= 0.50 &
                                                               x [SPINT1_id,  ]  >= 2.00 &
                                                               x [ELF5_id,    ]  >= 0.50 &
                                                               x [BCAM_id,    ]  >= 2.00 &
                                                               x [ERVFRD_id,  ]  <  1.00 &
                                                               x [SDC1_id,    ]  <  1.00 &
                                                               x [HLAG_id,    ]  <  2.00   })
MPI.monocle   <- classifyCells (MPI.monocle, MPI.monocle.2)
```

```{r, Figure 3A MPI (in vivo) analysis}
MPI_marker_diff              <- markerDiffTable   (MPI.monocle[invivo.monocle_expressed_genes, ], MPI.monocle.2, cores = 4)
MPI_semisup_clustering_genes <- row.names         (MPI_marker_diff)[order (MPI_marker_diff$qval)][1:500]
MPI.monocle                  <- setOrderingFilter (MPI.monocle, MPI_semisup_clustering_genes)

rm (MPI.monocle.2)
rm (MPI_marker_diff)
rm (MPI_semisup_clustering_genes)
```

```{r, Figure 3A MPI (in vivo) analysis}
MPI.monocle <- reduceDimension (MPI.monocle,
                                   max_components = 2,
                                   method         = 'DDRTree',
                                   norm_method    = 'none',
                                   pseudo_expr    = 0)
MPI.monocle <- orderCells      (MPI.monocle)
MPI.monocle <- orderCells      (MPI.monocle, root_state = 3)
```

```{r, Figure 3A MPI (in vivo) Monocle2 pseudotime trajectory}
plot_cell_trajectory (MPI.monocle,
                      color_by           =   "seurat_clusters",
                      cell_size          = 3,
                      show_branch_points = FALSE) +
scale_color_manual   (values             = c("#F8766D",
                                             "#CD9600",
                                             "#7CAE00",
                                             "#00BE67",
                                             "#C77CFF",
                                             "#00A9FF",
                                             "#00BFC4",
                                             "#FF61CC"),
                      name               =   "cluster") + NoLegend ()
```

```{r, Figure 3B MPI (in vivo) analysis}
MPI_BEAM_res <- BEAM (MPI.monocle, branch_point = 1, cores = 1)
MPI_BEAM_res <- MPI_BEAM_res [order (MPI_BEAM_res$qval), ]
MPI_BEAM_res <- MPI_BEAM_res [ , c("gene_short_name", "pval", "qval")]

head (MPI_BEAM_res)
```

```{r, Figure 3B MPI (in vivo) gene selection}
TARGETED_genes <- row.names (subset (fData (MPI.monocle), gene_short_name %in% c("BCAM",
                                                                                 "SPINT1",
                                                                                 "TP63",
                                                                                 "ELF5",
                                                                                 "TEAD4",
                                                                                 "ERVW-1",
                                                                                 "ERVFRD-1",
                                                                                 "CGB",
                                                                                 "NOTCH1",
                                                                                 "ITGA2",
                                                                                 "NOTCH2",
                                                                                 "ITGA5",
                                                                                 "HLA-G",
                                                                                 "ITGA1")))
```

```{r, Figure 3B MPI (in vivo) branched Monocle2 pseudotime heatmap}
MPI_branched_heatmap <- plot_genes_branched_heatmap (MPI.monocle[TARGETED_genes, ],
                                                     branch_point        = 1,
                                                     branch_labels       = c("EVT", "SCT"),
                                                     cores               = 1,
                                                     use_gene_short_name = TRUE,
                                                     show_rownames       = TRUE,
                                                     cluster_rows        = FALSE,
                                                     num_clusters        = 1, 
                                                     return_heatmap      = FALSE )
```

### PD-TOrg (in vitro) trophoblasts

```{r, Figure 3A PD-TOrg (in vitro) analysis}
PD.TOrg.monocle <- as.CellDataSet      (PD.TOrg, assay = "RNA")
PD.TOrg.monocle <- estimateSizeFactors (PD.TOrg.monocle)
PD.TOrg.monocle <- estimateDispersions (PD.TOrg.monocle)
PD.TOrg.monocle <- detectGenes         (PD.TOrg.monocle, min_expr = 0.1)
                   print               (head (fData (PD.TOrg.monocle)))
```

```{r, Figure 3A PD-TOrg (in vitro) analysis}
PD.TOrg.monocle_expressed_genes <- row.names (subset (fData (PD.TOrg.monocle), num_cells_expressed >= 10))
```

```{r, Figure 3A PD-TOrg (in vitro) analysis}
ERVFRD_id <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "ERVFRD-1"))
HLAG_id   <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "HLA-G"   ))
CDX2_id   <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "CDX2"    ))
BCAM_id   <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "BCAM"    ))
SPINT1_id <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "SPINT1"  ))
ELF5_id   <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "ELF5"    ))
SDC1_id   <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "SDC1"    ))
GCM1_id   <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name == "GCM1"    ))

PD.TOrg.monocle.2 <- newCellTypeHierarchy ()
PD.TOrg.monocle.2 <- addCellType   (PD.TOrg.monocle.2,
                                    "SCT",
                                    classify_func = function (x) { x [ERVFRD_id,  ]  >= 1.00 &
                                                                   x [GCM1_id,    ]  >= 2.00 &
                                                                   x [SDC1_id,    ]  >= 1.00   })
PD.TOrg.monocle.2 <- addCellType   (PD.TOrg.monocle.2,
                                    "EVT",
                                    classify_func = function (x) { x [HLAG_id,    ]  >= 2.00 &
                                                                   x [GCM1_id,    ]  >= 2.00 &
                                                                   x [SDC1_id,    ]  >= 1.00   })
PD.TOrg.monocle.2 <- addCellType   (PD.TOrg.monocle.2,
                                    "CTB",
                                    classify_func = function (x) { x [CDX2_id,    ]  >= 0.50 &
                                                                   x [SPINT1_id,  ]  >= 2.00 &
                                                                   x [ELF5_id,    ]  >= 0.50 &
                                                                   x [BCAM_id,    ]  >= 2.00 &
                                                                   x [ERVFRD_id,  ]  <  1.00 &
                                                                   x [SDC1_id,    ]  <  1.00 &
                                                                   x [HLAG_id,    ]  <  2.00   })
PD.TOrg.monocle   <- classifyCells (PD.TOrg.monocle, PD.TOrg.monocle.2)
```

```{r, Figure 3A PD-TOrg (in vitro) analysis}
PD.TOrg_marker_diff              <- markerDiffTable   (PD.TOrg.monocle[PD.TOrg.monocle_expressed_genes, ], PD.TOrg.monocle.2, cores = 4)
PD.TOrg_semisup_clustering_genes <- row.names         (PD.TOrg_marker_diff)[order (PD.TOrg_marker_diff$qval)][1:250]
PD.TOrg.monocle                  <- setOrderingFilter (PD.TOrg.monocle, PD.TOrg_semisup_clustering_genes)

rm (PD.TOrg.monocle.2)
rm (PD.TOrg_marker_diff)
rm (PD.TOrg_semisup_clustering_genes)
```

```{r, Figure 3A PD-TOrg (in vitro) analysis}
PD.TOrg.monocle <- reduceDimension (PD.TOrg.monocle,
                                   max_components = 2,
                                   method         = 'DDRTree',
                                   norm_method    = 'none',
                                   pseudo_expr    = 0)
PD.TOrg.monocle <- orderCells     (PD.TOrg.monocle)
PD.TOrg.monocle <- orderCells     (PD.TOrg.monocle, root_state = 3)
```

```{r, Figure 3A PD-TOrg (in vitro) Monocle2 pseudotime trajectory}
plot_cell_trajectory (PD.TOrg.monocle,
                      color_by           =   "seurat_clusters",
                      cell_size          = 3,
                      show_branch_points = FALSE) +
scale_color_manual   (values             = c("#F8766D",
                                             "#7CAE00",
                                             "#00A9FF",
                                             "#C77CFF",
                                             "#00BFC4",
                                             "#FF61CC"),
                      name               =   "cluster") + NoLegend ()
```

```{r, Figure 3B PD-TOrg (in vitro) analysis}
PD.TOrg_BEAM_res <- BEAM (PD.TOrg.monocle, branch_point = 1, cores = 4)
PD.TOrg_BEAM_res <- PD.TOrg_BEAM_res [order (PD.TOrg_BEAM_res$qval), ]
PD.TOrg_BEAM_res <- PD.TOrg_BEAM_res [ , c("gene_short_name", "pval", "qval")]

head (PD.TOrg_BEAM_res)
```

```{r, Figure 3B PD-TOrg (in vitro) gene selection}
TARGETED_genes <- row.names (subset (fData (PD.TOrg.monocle), gene_short_name %in% c("BCAM",
                                                                                     "SPINT1",
                                                                                     "TP63",
                                                                                     "ELF5",
                                                                                     "TEAD4",
                                                                                     "ERVW-1",
                                                                                     "ERVFRD-1",
                                                                                     "CGB",
                                                                                     "NOTCH1",
                                                                                     "ITGA2",
                                                                                     "NOTCH2",
                                                                                     "ITGA5",
                                                                                     "HLA-G",
                                                                                     "ITGA1")))
```

```{r, Figure 3B PD-TOrg (in vitro) branched Monocle2 pseudotime heatmap}
PD.TOrg_branched_heatmap <- plot_genes_branched_heatmap (PD.TOrg.monocle[TARGETED_genes, ],
                                                         branch_point        = 1,
                                                         branch_labels       = c("EVT", "SCT"),
                                                         cores               = 1,
                                                         use_gene_short_name = TRUE,
                                                         show_rownames       = TRUE,
                                                         cluster_rows        = FALSE,
                                                         num_clusters        = 1, 
                                                         return_heatmap      = FALSE )
```

### hTSC-TOrg (in vitro) trophoblasts

```{r, Figure 3A hTSC-TOrg (in vitro) analysis}
hTSC.TOrg.monocle <- as.CellDataSet      (hTSC.TOrg, assay = "RNA")
hTSC.TOrg.monocle <- estimateSizeFactors (hTSC.TOrg.monocle)
hTSC.TOrg.monocle <- estimateDispersions (hTSC.TOrg.monocle)
hTSC.TOrg.monocle <- detectGenes         (hTSC.TOrg.monocle, min_expr = 0.1)
                     print               (head (fData (hTSC.TOrg.monocle)))
```

```{r, Figure 3A hTSC-TOrg (in vitro) analysis}
hTSC.TOrg.monocle_expressed_genes <- row.names (subset (fData (hTSC.TOrg.monocle), num_cells_expressed >= 10))
```

```{r, Figure 3A hTSC-TOrg (in vitro) analysis}
ERVFRD_id <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name == "ERVFRD-1"))
HLAG_id   <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name == "HLA-G"   ))
BCAM_id   <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name == "BCAM"    ))
SPINT1_id <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name == "SPINT1"  ))
ELF5_id   <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name == "ELF5"    ))
SDC1_id   <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name == "SDC1"    ))
GCM1_id   <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name == "GCM1"    ))

hTSC.TOrg.monocle.2 <- newCellTypeHierarchy ()
hTSC.TOrg.monocle.2 <- addCellType   (hTSC.TOrg.monocle.2,
                                      "SCT",
                                      classify_func = function (x) { x [ERVFRD_id,  ]  >= 1.00 &
                                                                     x [GCM1_id,    ]  >= 2.00 &
                                                                     x [SDC1_id,    ]  >= 1.00   })
hTSC.TOrg.monocle.2 <- addCellType   (hTSC.TOrg.monocle.2,
                                      "EVT",
                                      classify_func = function (x) { x [HLAG_id,    ]  >= 2.00 &
                                                                     x [GCM1_id,    ]  >= 2.00 &
                                                                     x [SDC1_id,    ]  >= 1.00   })
hTSC.TOrg.monocle.2 <- addCellType   (hTSC.TOrg.monocle.2,
                                      "CTB",
                                      classify_func = function (x) { x [SPINT1_id,  ]  >= 2.00 &
                                                                     x [ELF5_id,    ]  >= 0.50 &
                                                                     x [BCAM_id,    ]  >= 2.00 &
                                                                     x [ERVFRD_id,  ]  <  1.00 &
                                                                     x [SDC1_id,    ]  <  1.00 &
                                                                     x [HLAG_id,    ]  <  2.00   })
hTSC.TOrg.monocle   <- classifyCells (hTSC.TOrg.monocle, hTSC.TOrg.monocle.2)
```

```{r, Figure 3A hTSC-TOrg (in vitro) analysis}
hTSC.TOrg_marker_diff              <- markerDiffTable   (hTSC.TOrg.monocle[hTSC.TOrg.monocle_expressed_genes, ], hTSC.TOrg.monocle.2, cores = 4)
hTSC.TOrg_semisup_clustering_genes <- row.names         (hTSC.TOrg_marker_diff)[order (hTSC.TOrg_marker_diff$qval)][1:500]
hTSC.TOrg.monocle                  <- setOrderingFilter (hTSC.TOrg.monocle, hTSC.TOrg_semisup_clustering_genes)

rm (hTSC.TOrg.monocle.2)
rm (hTSC.TOrg_marker_diff)
rm (hTSC.TOrg_semisup_clustering_genes)
```

```{r, Figure 3A hTSC-TOrg (in vitro) analysis}
hTSC.TOrg.monocle <- reduceDimension (hTSC.TOrg.monocle,
                                      max_components = 2,
                                      method         = 'DDRTree',
                                      norm_method    = 'none',
                                      pseudo_expr    = 0)
hTSC.TOrg.monocle <- orderCells      (hTSC.TOrg.monocle)
hTSC.TOrg.monocle <- orderCells      (hTSC.TOrg.monocle, root_state = 3)
```

```{r, Figure 3A hTSC-TOrg (in vitro) Monocle2 pseudotime trajectory}
plot_cell_trajectory (hTSC.TOrg.monocle,
                      color_by           =   "seurat_clusters",
                      cell_size          = 3,
                      show_branch_points = FALSE) +
scale_color_manual   (values             = c("#7CAE00",
                                             "#00A9FF",
                                             "#00BFC4",
                                             "#F8766D",
                                             "#AEA200",
                                             "#FF61CC",
                                             "#C77CFF"),
                      name               =   "cluster") + NoLegend ()
```

```{r, Figure 3B hTSC-TOrg (in vitro) analysis}
invitro_BEAM_res <- BEAM (hTSC.TOrg.monocle, branch_point = 1, cores = 4)
invitro_BEAM_res <- invitro_BEAM_res [order (invitro_BEAM_res$qval), ]
invitro_BEAM_res <- invitro_BEAM_res [ , c("gene_short_name", "pval", "qval")]

head (invitro_BEAM_res)
```

```{r, Figure 3B hTSC-TOrg (in vitro) gene selection}
TARGETED_genes <- row.names (subset (fData (hTSC.TOrg.monocle), gene_short_name %in% c("BCAM",
                                                                                       "SPINT1",
                                                                                       "TP63",
                                                                                       "ELF5",
                                                                                       "TEAD4",
                                                                                       "ERVW-1",
                                                                                       "ERVFRD-1",
                                                                                       "CGB",
                                                                                       "NOTCH1",
                                                                                       "ITGA2",
                                                                                       "NOTCH2",
                                                                                       "ITGA5",
                                                                                       "HLA-G",
                                                                                       "ITGA1")))
```

```{r, Figure 3B hTSC-TOrg (in vitro) branched Monocle2 pseudotime heatmap}
hTSC.TOrg_branched_heatmap  <- plot_genes_branched_heatmap (hTSC.TOrg.monocle[TARGETED_genes, ],
                                                            branch_point        = 1,
                                                            branch_labels       = c("EVT", "SCT"),
                                                            cores               = 1,
                                                            use_gene_short_name = TRUE,
                                                            show_rownames       = TRUE,
                                                            cluster_rows        = FALSE,
                                                            num_clusters        = 1, 
                                                            return_heatmap      = FALSE )
```

## Monocle3 pseudotime analyses

### MPI (in vivo) trophoblasts

```{r, Figure 3C MPI analysis}
MPI.monocle3 <- as.cell_data_set      (MPI, assay = "RNA")
MPI.monocle3 <- estimate_size_factors (MPI.monocle3)
MPI.monocle3 <- cluster_cells         (MPI.monocle3,
                                       reduction_method = c("UMAP"),
                                       cluster_method   = "leiden",
                                       partition_qval   = 0.05,
                                       resolution       = 0.0005)

p1  <- plot_cells (MPI.monocle3,                               show_trajectory_graph = FALSE)
p2  <- plot_cells (MPI.monocle3, color_cells_by = "partition", show_trajectory_graph = FALSE)
wrap_plots        (p1, p2)
```

```{r, Figure 3C MPI analysis}
MPI.monocle3 <- learn_graph (MPI.monocle3, use_partition = FALSE, close_loop = TRUE)
```

```{r, Figure 3C MPI pseudotime UMAP}
MPI.monocle3 <- order_cells (MPI.monocle3)

plot_cells (MPI.monocle3,
            color_cells_by                = "pseudotime",
            label_cell_groups             = FALSE,
            label_leaves                  = FALSE,
            label_branch_points           = FALSE,
            graph_label_size              = 1.5,
            cell_size                     = 3,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25)
```

### PD-TOrg (in vitro) trophoblasts

```{r, Figure 3C PD-TOrg analysis}
PD.TOrg.monocle3 <- as.cell_data_set      (exvivo, assay = "RNA")
PD.TOrg.monocle3 <- estimate_size_factors (PD.TOrg.monocle3)
PD.TOrg.monocle3 <- cluster_cells         (PD.TOrg.monocle3,
                                           reduction_method = c("UMAP"),
                                           cluster_method   = "leiden",
                                           partition_qval   = 0.05,
                                           resolution       = 0.0005)

p1  <- plot_cells (PD.TOrg.monocle3,                               show_trajectory_graph = FALSE)
p2  <- plot_cells (PD.TOrg.monocle3, color_cells_by = "partition", show_trajectory_graph = FALSE)
wrap_plots        (p1, p2)
```

```{r, Figure 3C PD-TOrg analysis}
PD.TOrg.monocle3 <- learn_graph (PD.TOrg.monocle3, use_partition = FALSE, close_loop = TRUE)
```

```{r, Figure 3C PD-TOrg pseudotime UMAP}
PD.TOrg.monocle3 <- order_cells (PD.TOrg.monocle3)

plot_cells (PD.TOrg.monocle3,
            color_cells_by                = "pseudotime",
            label_cell_groups             = FALSE,
            label_leaves                  = FALSE,
            label_branch_points           = FALSE,
            graph_label_size              = 1.5,
            cell_size                     = 3,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25)
```

### hTSC-TOrg (in vitro) trophoblasts

```{r, Figure 3C hTSC-TOrg analysis}
hTSC.TOrg.monocle3 <- as.cell_data_set      (invitro, assay = "RNA")
hTSC.TOrg.monocle3 <- estimate_size_factors (hTSC.TOrg.monocle3)
hTSC.TOrg.monocle3 <- cluster_cells         (hTSC.TOrg.monocle3,
                                             reduction_method = c("UMAP"),
                                             cluster_method   = "leiden",
                                             partition_qval   = 0.05,
                                             resolution       = 0.0015)

p1  <- plot_cells (hTSC.TOrg.monocle3,                               show_trajectory_graph = FALSE)
p2  <- plot_cells (hTSC.TOrg.monocle3, color_cells_by = "partition", show_trajectory_graph = FALSE)
wrap_plots        (p1, p2)
```

```{r, Figure 3C hTSC-TOrg analysis}
hTSC.TOrg.monocle3 <- learn_graph (hTSC.TOrg.monocle3, use_partition = FALSE, close_loop = TRUE)
```

```{r, Figure 3C hTSC-TOrg pseudotime UMAP}
hTSC.TOrg.monocle3 <- order_cells (hTSC.TOrg.monocle3)

plot_cells (hTSC.TOrg.monocle3,
            color_cells_by                = "pseudotime",
            label_cell_groups             = FALSE,
            label_leaves                  = FALSE,
            label_branch_points           = FALSE,
            graph_label_size              = 1.5,
            cell_size                     = 3,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25)
```

## Model-speicfic origin comparison

```{r, Labelling the model-specific origins}
MPI_ORIGIN <- AddMetaData (MPI_ORIGIN, metadata = "MPI", col.name = "ORIGIN")
MPI_ORIGIN[[]]

PD.TOrg_ORIGIN <- AddMetaData (PD.TOrg_ORIGIN, metadata = "PD.TOrg", col.name = "ORIGIN")
PD.TOrg_ORIGIN[[]]

hTSC.TOrg_ORIGIN <- AddMetaData (hTSC.TOrg_ORIGIN, metadata = "hTSC.TOrg", col.name = "ORIGIN")
hTSC.TOrg_ORIGIN[[]]
```

```{r, Merging all model-specific origins}
ORIGIN <- merge(x            = MPI_ORIGIN,
                y            = c(PD.TOrg_ORIGIN, hTSC.TOrg_ORIGIN),
                add.cell.ids = NULL,
                merge.data   = TRUE)
ORIGIN[[]]
```

# PCC Comparison

```{r, Figure 3D analysis}
Idents (ORIGIN) <- ORIGIN@meta.data$ORIGIN
table (Idents (ORIGIN))

av.exp    <- AverageExpression (ORIGIN)$RNA
cor.exp   <- as.data.frame     (cor (av.exp))
cor.exp$x <- rownames          (cor.exp)
cor.df    <- tidyr::gather     (data = cor.exp, y, correlation, c('MPI', 'PD.TOrg', 'hTSC.TOrg'))
cor.df
```

```{r, Figure 3D}
ggplot               (cor.df, aes (x, y, fill = correlation)) +
geom_tile            () +
scale_fill_gradientn (limits = c(0,1), colours = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100))
```

# Differential Expression

```{r, Merging specific model-specific origins}
ORIGIN2 <- merge(x           = MPI_ORIGIN,
                y            = PD.TOrg_ORIGIN,
                add.cell.ids = NULL,
                merge.data   = TRUE)
ORIGIN2[[]]

ORIGIN3 <- merge(x            = MPI_ORIGIN,
                 y            = hTSC.TOrg_ORIGIN,
                 add.cell.ids = NULL,
                 merge.data   = TRUE)
ORIGIN3[[]]
```

```{r, Figure 3E MPI vs PD-TOrg origin analysis}
DefaultAssay (ORIGIN2) <- "RNA"

Idents (ORIGIN2) <- ORIGIN@meta.data$ORIGIN2
ORIGIN.response  <- FindMarkers (ORIGIN,
                                 ident.1            = "MPI",
                                 ident.2            = "PD.TOrg",
                                 test.use           = "MAST",
                                 logfc.threshold    = -Inf,
                                 min.pct            = -Inf,
                                 min.diff.pct       = -Inf,
                                 verbose            = FALSE)
                    head        (ORIGIN.response, n = 50)
```

```{r, Table S6A}
write.xlsx (ORIGIN.response, file = "MPI_vs_PD-TOrg_originDGE.xlsx")
```

```{r, Figure 3E MPI vs PD-TOrg origin volcano plot}
EnhancedVolcano (ORIGIN.response,
                 lab             = rownames (ORIGIN.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 xlim            = c(min (ORIGIN2.response[["avg_log2FC"]], na.rm = TRUE) - 2.5,
                                     max (ORIGIN2.response[["avg_log2FC"]], na.rm = TRUE) + 1.5),
                 gridlines.major = FALSE,
                 selectLab       = c('CGA', 'NEAT1', 'BCAM', 'PAGE4', 'EFEMP1', 'PEG10', 'EGR1'),
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
```

```{r, Figure 3E MPI vs hTSC-TOrg origin analysis}
DefaultAssay (ORIGIN2) <- "RNA"

Idents (ORIGIN2) <- ORIGIN2@meta.data$ORIGIN
ORIGIN2.response <- FindMarkers (ORIGIN2,
                                 ident.1            = "MPI",
                                 ident.2            = "hTSC.TOrg",
                                 test.use           = "MAST",
                                 logfc.threshold    = -Inf,
                                 min.pct            = -Inf,
                                 min.diff.pct       = -Inf,
                                 verbose            = FALSE)
                    head        (ORIGIN2.response, n = 50)
```

```{r, Table S6B}
write.xlsx (ORIGIN2.response, file = "MPI_vs_hTSC-TOrg_originDGE.xlsx")
```

```{r, Figure 3E MPI vs hTSC-TOrg origin volcano plot}
EnhancedVolcano (ORIGIN2.response,
                 lab             = rownames (ORIGIN2.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 xlim            = c(min (ORIGIN2.response[["avg_log2FC"]], na.rm = TRUE) - 2.5,
                                     max (ORIGIN2.response[["avg_log2FC"]], na.rm = TRUE) + 1.5),
                 gridlines.major = FALSE,
                 selectLab       = c('FTL', 'CGA', 'CD9', 'NEAT1', 'HLA-G', 'ITGA2', 'BCAM', 'EFEMP1', 'PEG10', 'EGR1', 'EPCAM', 'FOXO4'),
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
```